---
title: "Extract NDVI"
author: "Aidan Brushett"
date: "2025-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, eval = FALSE}
library(remotes)
install_github("ropensci/MODIStsp")
```

```{r}
library(sf) # for points and buffering
library(tidyverse)
library(terra)
#library(MODIStsp)
rm(list = ls())
```

# 1. Read in the raw site coordinates:

We want these separated so I won't bind them with purrr just yet. Clunky but versatile. 

```{r}
sites_osm <- read_csv("./data/input/OSM_site_coordinates_2021-2024.csv") %>%
  
  st_as_sf(., coords = c("easting_12n", "northing_12n"), crs = 26912)


# Sampled 2022 onwards
#sites_osm_2022 <- sites_osm %>% 
#  
#  filter(array %in% c("LU13", "LU15", "LU21", "LU1",
#                      "LU9", "LU16", "LU14", "LU22"))
#
# Sampled 2021 
#sites_osm_2021 <- sites_osm %>% 
#  
#  filter(array %in% c("LU2", "LU3"))






# buffer sites so that we can filter the landcover data and reduce file size
sites_osm_buffered <- st_buffer(sites_osm, dist = 11000) %>%
  
  st_union()

st_bbox(st_transform(sites_osm_buffered, crs = 4326))


plot(sites_osm_buffered)

st_write(sites_osm_buffered, "./data/input/modis_sites.shp", append = FALSE)

```

# 2. Read in the NDVI

The NDVI layer used here was generated using the following javascript code in **Google Earth Engine**:

```{javascript, eval = FALSE}
var aoi = ee.Geometry.Rectangle([-115.22791, 54.46367, -109.91320, 57.65839]);
Map.centerObject(aoi, 6);  // Zoom level 6 is usually good for this scale
Map.addLayer(aoi, {color: 'red'}, 'AOI');


// 2. Filter MODIS NDVI (MOD13A1 = 16-day NDVI, 500m resolution)
var dataset = ee.ImageCollection('MODIS/061/MOD13A1')
                  .filterDate('2020-05-15', '2024-09-15')
                  .filter(ee.Filter.calendarRange(5, 9, 'month'))  // May to September
                  .select('NDVI');

// Select the NDVI collection from the dataset then clip to the area of interest
var ndvi = dataset.select('NDVI').map(function(image) {
  return image.clip(aoi);
});

//Take the mean of the NDVI image collection
var ndvi_mean = ndvi.mean();

// Set some color schemes and parameters for layers
var ndviVis = {
  min: 0,
  max: 9000,
  palette: [
    'ffffff', 'ce7e45', 'df923d', 'f1b555', 'fcd163', '99b718', '74a901',
    '66a000', '529400', '3e8601', '207401', '056201', '004c00', '023b01',
    '012e01', '011d01', '011301'
  ],
};

//Add to map
Map.addLayer(ndvi_mean, ndviVis, 'NDVI Mean');
Map.addLayer(ndvi, ndviVis, 'NDVI');



//Export layer
Export.image.toDrive({
  image: ndvi_mean,
  description: 'NDVI_Mean_Export',
  folder: 'EarthEngineExports',          // Optional: change to desired folder
  fileNamePrefix: 'NDVI_Mean_2020_2024', // Optional: change file name
  region: aoi,                           // Your area of interest
  scale: 500,                            // MODIS native resolution
  maxPixels: 1e13
});
```

Let's read in the raster that came from that script:

```{r}
# Modis stores integers that need to be re-scaled for the actual NDVI values

ndvi_raw <- rast("data/input/NDVI_Mean_2020_2024.tif") * 0.0001

plot(ndvi_raw)
```

Let's mask out the NDVI values that are below 0 (i.e., not vegetated):

```{r}
ndvi <- classify(ndvi_raw, rcl = matrix(c(-Inf, 0.1, NA), ncol=3, byrow=TRUE))

plot(ndvi)
```

# 3. Extract NDVI 

Currently set to suppress warnings about transforming the vector data to the CRS of the raster. 

```{r warning=FALSE}
sites <- sites_osm
  
# Buffer distances for which we are extracting using purrr
sites_ndvi <- purrr::map_dfr(c(50, 100, 200, 250, 300, 400, seq(500, 5000, by = 250), 10000), ~{
  
  st_buffer(sites, .x) %>%
    
    extract(ndvi, ., fun = mean, na.rm = TRUE) %>%
    
    bind_cols(sites, .) %>%
    
    mutate(buffer_dist = .x) %>%
      
    relocate(buffer_dist, .after = lat) %>%
      
    st_drop_geometry(.)
  
  }) %>%
  
  rename(ndvi = NDVI) %>%
  
  # Clean up
  select(-any_of(c("ID", "id", "long", "lat", "easting_12n", "northing_12n", "start_date", 
                   "end_date", "camera_failure_details", "deployment_id" )))
#MODIStsp()
```

Save the file:

```{r}
write_csv(sites_ndvi, file="./data/output/OSM_MODISNDVI_site_covariates.csv")
```



