---
title: "ACME Timelapse Image Processing and Error Checking"
author: "Aidan Brushett"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: default
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
```

# 0. Before you begin

## 0.1. Notes

A few notes about this script. The Markdown version is set to eval=FALSE for ALL chunks to help with knitting speed. This script is modular and it's meant to be run in sections. It could be considered more of a toolkit/workbook then a top-to-bottom script. Unless your dataset is 100% clean, it is unlikely that you will ever run it top to bottom. Remove the global `opts_chunk$set` settings if you would like to see script outputs in the knitted document. 

If you are running this with the 2024-2024 data make sure you download the whole (OSM_2024-2025 GitHub repository)[https://github.com/ACMElabUvic/OSM_2024-2025] from the ACMElabUvic GitHub. This will ensure you have all the files, data, and proper folder structure you will need to run this code and associated analyses.


## Netdrive access

This script relies on the user having access to the ACME lab Netdrive (you can view the .html output of this file if you don't have access and just want to see what the script did).

Helpful instructions for connecting to and navigating the Netdrive can also be found here: [https://docs.google.com/document/d/1Z72IrlIXO8MUHCoVztcMrMdL10R2tHrHThEgfow1Cu0/edit ](https://docs.google.com/document/d/1Z72IrlIXO8MUHCoVztcMrMdL10R2tHrHThEgfow1Cu0/edit).

***

# 1. Set up workspace

## 1.1. Install packages

If you don't already have the following packages installed, use the code below to install them. *NOTE this will not run automatically as eval=FALSE is included in the chunk setup (i.e. I don't want it to run everytime I run this code since I have the packages installed)

```{r install, eval=FALSE}
install.packages("tidyverse") 
install.packages("withr") 
install.packages("writexl")
install.packages("RSQLite")
install.packages("fs")
install.packages("crayon")
install.packages("magick")
```

## 1.2. Load libraries

And a couple custom functions that can help with masking issues from package loading:

```{r libraries, message=FALSE, warning=FALSE}
rm(list=ls()) # clear environment


library(tidyverse) # data tidying, visualization, and much more; this will load all tidyverse packages, can see complete list using tidyverse_packages()
library(withr) # used to temporarily set wd
library(RSQLite) # used to access timelapse data stored in ddb files
library(ggplot2) # pretty plots
library(crayon) # custom color for warning messages. Aesthetic only.
library(writexl) # writing error check spreadsheets as an excel sheet so formatting sticks
library(fs) # for writing and copying files and directories
library(magick) # for blurring or manipulating images

# These are functions that can sometimes get masked by other packages. Shouldn't be an issue and could be removed.
select <- dplyr::select
filter <- dplyr::filter
summarize <- dplyr::summarize
# Cheeky function that means "not in"
`%nin%` = Negate(`%in%`)

```

## 1.3. Script parameters

The ACME netdrive location varies between Apple/Windows users, and between Windows users depending on how they mapped their drive. Specify that (as well as some other parameters) here before running the script. 

```{r script_parameters}
my_netdrive <- getwd() # the path to the netdrive on your device

# Do you want to automatically append a blank noon photo (with snow) at sites where noon photos are missing? See Section 5.  
autofix_one_image_per_day <- FALSE
interpolate_snow <- FALSE

# Do you want to blur human use images at the end of the script? This step is currently not used and stored in the Appendix. 
blur_human_use <- FALSE
```

***





# 3. Import Timelapse data

## 3.1. Identify *.ddb* files to be imported. 

First, we need to temporarily set your working drive to the ACME lab Netdrive to import the Timelapse data files (there are too many to efficiently store on GitHub or GoogleDrive each year). We will use the `with_dir()` function in the *withr* package to do this.

**As of 2024: .csv files are not used for any step of data importing or processing to prevent user errors.** The Timelapse2 *.ddb* files are actually just SQLite databases with a different file extension! Anything you tag is automatically stored in the .ddb file's *datatable*, which we will access later to retrieve image data. The code below will identify all the files with a .ddb extension from any subfolder/s within the main folder provided in the `path =` argument. I have also specified for it to ignore certain subfolders (e.g. backups). Then it will name each file based on the base name of the .ddb file.

This only works to import the timelapse .ddb files **if** they are found within the subfolders of the main path folder, and there aren't extraneous files with a .ddb extension that we don't want to import. **Ensure nobody has a Quickpaste file stored in *1. Imagery*, with the exception of the *1. Imagery/Quickpaste/* directory.**
If the Netdrive folder structure changes in future years this code *may* not work or *may* need to be adapted to ignore additional subfolders.

```{r find, eval=TRUE}
# temporarily set the working directory to import from the NetDrive
with_dir(new = my_netdrive,
         
# make a list of the .csv file names from the sub-folders within the '1. Imagery/OSM_2022_LU01' folder
OSM_2024_timelapse_ddb_files <- list.files(
  #path = '1. Imagery',
  recursive = TRUE,
  pattern = "\\.*ddb$",
  full.names = TRUE
  ) %>% 

  
  # exclude files from backups sub-folders
  stringr::str_subset(.,
                      'Backups',
                      negate = TRUE) %>%
  # and exclude anyone's quickpaste files
  stringr::str_subset(.,
                      'Quickpaste',
                      negate = TRUE)
)

# Save the list. This way, you do not need to repeat this step every time you re-check for errors. 
save(OSM_2024_timelapse_ddb_files, file="./OSM_2024_timelapse_ddb_files.Rdata")
```

> This takes a while to run because R is looking in every sub-folder at every file to see if it matches the extension (i.e. .csv) and then saving the file name and path and then going back and removing the files from the 'backups' sub-folders. Maybe not the most elegant way to do this but beats having to manually copy and paste each file individually.

## 3.2. If running the script in steps: re-import the .ddb source list

Remember: this is only a SNAPSHOT of the .ddb files in the Imagery folder, and could be outdated if it's been a while since you ran the chunk above this one (e.g., somebody classifies a new folder). Make sure you re-run the `list.files()` function in the chunk above if you suspect there are new Timelapse files, or if their directory has changed. 
```{r re-import ddb list, eval=TRUE}
load("./OSM_2024_timelapse_ddb_files.Rdata")
```

## 3.3. Custom function to import .ddb files

This function is written to perform several key tasks. **As of 2024, Start/End tags for *event* are added automatically to prevent user errors.**

- imports and formats the 'datatable' in each .ddb file.  
- automatically applies the Start/End tags.
- attempts to standardize site names by converting dashes to underscores, and removing leading zeroes. If site names are particularly f**ed up, they might need manual fixing. 

All the .ddb files are named identically during tagging; throughout this script we will keep track of the data source of each Timelapse entry using the *datasource* and *fullpath* columns. 

> This code chunk only defines the function used to import the data-- no importing is actually performed in this step.

```{r import ddb files - function ONLY}
process_ddb_file <- function(ddb_file) {
  
  # connect to the .ddb file
  db <- dbConnect(RSQLite::SQLite(), dbname = ddb_file)
  
  # extract the datatable from the .ddb database. The internal structure of these is quite cool. 
  ddb_raw <- dbReadTable(db, "DataTable") %>%
    
    as_tibble() %>%
    
    # specify how you want to read in each column.
    type_convert(., col_types = cols(RootFolder = col_character(),
                                     File = col_character(),
                                     RelativePath = col_character(),
                                     Dark = col_logical(),
                                     DeleteFlag = col_logical(),
                                     Site = col_factor(),
                                     Classifier = col_factor(),
                                     Snow = col_factor(),
                                     Species = col_factor(),
                                     Event = col_character(),
                                     Empty = col_logical(),
                                     CoatColour = col_character(),
                                     CameraMalfunction = col_factor(),
                                     OtherSpecify = col_character(),
                                     Comments = col_character(),
                                     Noteworthy = col_logical(),
                                     DateTime = col_datetime(format = "%Y-%m-%d %H:%M:%S"),
                                     LeftAntler = col_factor(),
                                     RightAntler = col_factor(),
                                     LCount = col_factor(),
                                     RCount = col_factor(),
                                     .default = col_integer()),
                 trim_ws = TRUE # timelapse sometimes adds leading spaces, especially to the datetime column. We want these gone from everything. 
                 ) %>%
    
    # Lowercase names
    set_names(names(.) %>%
                tolower()) %>%
    
    # Create the fullpath and datasource columns, which we will use to keep track of where image data originally came from. 
    mutate(fullpath = paste0(dirname(ddb_file), "/",
                             ifelse(is.na(relativepath), "", paste0(relativepath, "/")), 
                             file),
           datasource = ddb_file
           ) %>%
    

    # now that we've done some basic checks, automatically apply the "Start" and "End" tags. 
    mutate(
        DateTime2 = as.POSIXct(datetime), # convert to seconds. Might be redundant code, this function was embedded from previous code written by Aidan.
        TimeDiff = c(NA, difftime(DateTime2[-1], DateTime2[-n()], units = "secs")), # time difference from prev photo
        episode_id = cumsum(ifelse(is.na(TimeDiff) | TimeDiff > 60,  # Create episode IDs to distinguish from another. 
                                  1, 
                                  0))
        ) %>%          
      
    group_by(episode_id) %>%
      
    mutate( 
      episodelength = n(), # how many images in the episode?
      event = ifelse(episodelength==1, 
                     "", # do not populate "Start" for singletons
                     case_when(  # Assign start and end to the first/last image of each episode, respectively
                       row_number() == 1 ~ "Start",
                       row_number() == n() ~ "End",
                       TRUE ~ ""))
      ) %>%
    
    ungroup() %>%
    
    select(-any_of(c('DateTime2', 'TimeDiff', 'dark'))) %>% # clean up
  
    # Make a first pass at cleaning up the site names. Probably a neater way of doing this, sue me. 
    mutate(site = gsub("-", "_", site) %>% # First, replace the hyphen with underscore
             gsub("_0+", "_", .) %>% # Then, remove leading zeros after the underscore
             gsub(" ", "", .) %>% # Then, remove any spaces
             as.factor(.) # Then, convert back to factor
           )  
  
  # Disconnect from the .ddb file. It is important to close the connection to avoid corrupting the data. 
  dbDisconnect(db)
  
  # CRITICAL errors that need to be fixed before running the script
    # Possible corrupted .ddb files. 
  if (nrow(ddb_raw) == 0) {
    stop(cat('\n The file \n', ddb_file, '\n contains no rows of data. The file may be corrupt. Please verify this in Timelapse before proceeding.'))
  }
  
  # Incorrect datetime format (common if Timelapse is set up incorrectly)
  if ("datetime" %nin% colnames(ddb_raw)) {
    stop(cat('\n The file \n', ddb_file, '\n may have an incorrect date time format. Please fix this in Timelapse before proceeding.'))
  }
  
  # Additional errors that should be resolved in Timelapse with the script re-run afterwards. 
  # `red()` here and elsewhere comes from the crayon package and changes the color of messages in the console. Can be removed if the package ever becomes deprecated. 
  
  # NA site values
  if (any(is.na(ddb_raw$site))) {
    warning(paste0('The',
                   crayon::red(' Site '), 
                   'column contains NA values in the file: \n  ', 
                   crayon::italic(ddb_file), '\n\n'))  }
  # Duplicate site values
  if(n_distinct(ddb_raw$site, na.rm=T) > 1) {
    warning(paste0('The',
                   red(' Site '), 
                   'column contains multiple values: \n  ', 
                   italic(ddb_file), '\n\n'))  }
  
  # NA classifier values
  if (any(is.na(ddb_raw$classifier))) {
    warning(paste0('The',
                   red(' Classifier '), 
                   'column contains NA values in the file: \n  ', 
                   italic(ddb_file), '\n\n'))  }

  # Multiple classifier values
  if(n_distinct(ddb_raw$classifier, na.rm=T) > 1) {
    warning(paste0('The',
                   red(' Classifier '), 
                   'column contains multiple values: \n  ', 
                   italic(ddb_file), '\n\n'))  }
  
  # Missing snow values
  if (any(is.na(ddb_raw$snow))) {
    warning(paste0('The',
                   red(' Snow '), 
                   'column is entirely NA values in the file: \n  ', 
                   italic(ddb_file), '\n\n'))  }
  
  # The end goal of the function: a tibble with the semi-cleaned, semi-formatted data from each site!!!
  return(ddb_raw)
    
} # end function
```

## 3.4. Read in, format, and append the .ddb files.
The function `purrr::map_dfr()` neatly appends each site's data into a single dataframe. This will import data from any .ddb file that is in the `OSM-2024_timelapse_ddb_files` object we created earlier. It handles missing columns nicely by filling in 'NA' for any sites missing those columns. The *dark* and *deleteflag* columns in particular tend to be in some files depending on the version of Timelapse used.

> Error messages need to be fixed immediately. Warning messages are 'FYI' only about high-level tagging errors, to be addressed later.

```{r import ddb files}
with_dir(new = my_netdrive,
  OSM_2024_data <- purrr::map_dfr(OSM_2024_timelapse_ddb_files, # the list of ddb files from earlier that we are importing. 
                                  process_ddb_file) # apply the function
)

# Export the data as a .csv. This will never be used again but it's always nice to have backups ;)
# write_csv(OSM_2024_data, file="./data/raw/OSM_2024_raw_import.csv")
```

***

# 4. Error checking Timelapse data.

The following sub-section check the Timelapse data that we just imported for classification errors. These are rule-based and include anything from missing snow values to mismatches between the deployment dates and the actual iamge dates in Timelapse. Each sub-section searches for a specific 'category' of error and performs the following:

- Uses a set of criteria based on the ACME protocol to identify a specific tagging error. Each chunk represents a specific error. 
- Binds together the specific error information into a single dataframe
- Exports the dataframe to Excel in the *./data/processed/* directory

After running a chunk, navigate to *./data/processed/* and use the spreadsheets to correct errors in Timelapse directly. Some are straightforward, others may take some sleuthing to figure out. Depending on the number of errors, this could be a group task and you can upload the spreadsheets to Google Drive. 

The 'categories' of errors are sorted in order from broad to specific. You may find that after fixing a metadata error, many image or datetime errors disappear on their own. For this reason, I recommend doing a first pass of each error 'category' before moving on to the next one. 

*This spreadsheet DOES NOT check the following:*
*- image timestamps. These could be wrong for many reasons and should be identified using slate photos and datasheets while tagging*
*- the accuracy of your species identification*
*- the accuracy of coat color, antler point, or demographic fields* 
*Remember: nothing replaces an attentive eye and careful detail while tagging!*

## 4.1. Duplicate data import errors

Self-explanatory. Steps to resolve errors include:

1. Open the spreadsheet with the error information
2. Find the .ddb files that have been duplicated
3. Determine if these are copied files? Old quickpaste files? Correct as needed.
4. Re-run Section 3.1 onward to generate an updated list of the .ddb files to be imported. 

```{r error duplicate}
OSM_2024_error_duplicate <- OSM_2024_data %>%
  
  group_by(across(everything())) %>%  # Group by all columns
  
  filter(n() > 1) %>%                # Keep groups with more than 1 row
  
  ungroup() %>%
  
  select(site, file, fullpath, datasource) %>% # clean up the dataframe
  
  distinct() %>%
  
  mutate(error = "Duplicate data for these entries") # custom error message


if(nrow(OSM_2024_error_duplicate) == 0){
  writeLines("Lucky you! No duplicate rows in the dataset.")
  rm(OSM_2024_error_duplicate)
  
} else {
  
  writexl::write_xlsx(OSM_2024_error_duplicate, "./OSM_2024_error_duplicate.xlsx", format_headers = TRUE)
  writeLines("Your dataset may have duplicate entries. Please refer to the excel sheet and correct in Timelapse directly")
}
```

## 4.3. Metadata/site info errors

These relate to the "heading" columns in Timelapse: *site*, *snow*, and *classifier.*. This is also where we will ensure that the site names in the data match the site names in the deployments (and vice versa). Note: if a camera was not retrieved during the field season (e.g. due to a road flooding), this will falsely be flagged as a 'datamissing' error. These issues should be marked in the camera_failure_details column of the deployment spreadsheet and can be distinguished from truly missing data. These can be disregarded at your discretion. 

Steps to resolve these include:

1. Open the spreadsheet with the error information.
2. Filter the spreadsheet to a specific site to view all errors at that camera site.
3. Correct mistakes directly in Timelapse, or in the deployment spreadsheet if necessary. 
4. Re-run Section 3.4. onwards to update the list of errors. 

Identify sites with missing values for snow:
```{r error snow}
error_snow <- OSM_2024_data %>%
  
  group_by(datasource) %>% # group by import
  
  filter(any(is.na(snow))) %>% # any NA values?
  
  select(site, datasource) %>% # clean up the dataframe
  
  distinct() %>%
  
  mutate(error = "NAs present in the snow column")
```

Identify sites with more than one value for classifier (or NAs):

```{r}
error_classifier <- OSM_2024_data %>%
  
  group_by(datasource) %>%
    
  filter(
    any(is.na(classifier)) | # At least one NA values
      all(is.na(classifier)) | # OR, all NA values (this code is redundant and could be deleted)
      n_distinct(classifier, na.rm=T) > 1 # OR, there are multiple non-NA classifiers entered. 
      ) %>%
  
  select(site, datasource) %>%
  
  distinct() %>%
  
  mutate(error = "NAs or multiple values present in the classifier column")
```

Identify sites with multiple (or missing) site names: 
```{r}
error_site <- OSM_2024_data %>%
  
  group_by(datasource) %>%
    
  filter(
    any(is.na(site)) | # At least one NA value
      all(is.na(site)) |  # OR, all NA values (this code is redundant and could be deleted)
      n_distinct(site, na.rm=T) > 1 # OR, there are multiple non-NA values
      ) %>%
  
  select(site, datasource) %>%
  
  distinct() %>%
  
  mutate(error = "NAs or multiple values present in the site column")
```


Combine all the errors into a single spreadsheet and export:
```{r}
# bind_rows neatly handles any columns that aren't present in every spreadsheet. Much better than rbind(). 
OSM_2024_error_metadata <- bind_rows(error_classifier, 
                                      error_snow, 
                                      error_site)

if(nrow(OSM_2024_error_metadata) == 0){
  writeLines("Lucky you! No metadata errors in the dataset.")
} else {
  writexl::write_xlsx(OSM_2024_error_metadata, "./OSM_2024_error_metadata.xlsx", format_headers = TRUE)
  writeLines("Metadata errors are present. Please refer to the excel sheet and correct in Timelapse directly if needed.")
}

rm(error_classifier, error_snow, error_site) # clean up
```

## 4.3. Species/image tagging errors

These errors are anything related to ACME's tagging protocol. These can be time-consuming to fix Missing coat color, missing species tags, incorrect demographics, etc. Note: when two animals occur together in an image, or within 60 seconds of each other, this will be falsely tagged as an error. These instances should be mentioned in the comments in Timelapse so they are easy to distinguish from true errors. They will *always* get flagged as errors. These can be disregarded at your discretion. 

Steps to resolve these include:

1. Open the spreadsheet with the error information
2. Filter the spreadsheet to a specific site to view all errors at that camera site
3. Correct mistakes directly in Timelapse. 
4. Re-run Section 3.4. onwards to update the list of errors.

**As of 2024, .csv spreadsheets are a thing of the past and DO NOT need to be re-exported after fixing errors in Timelapse.** Everything is updated in real time in the .ddb files, woohoo!

Missing or multiple species tags for a non-empty event:
```{r}
error_speciestag <- OSM_2024_data %>%
  
  group_by(site, episode_id, datasource) %>% # Including datasource as a grouping variable is necessary when there are >9999 images per site. retrieve every image in an event
  
  filter(any(empty==FALSE)) %>% # Select non-empty events

  filter(
    n_distinct(species)!=1 # Select events with more than one species tag (this includes NA)
    ) %>%
  
  mutate(
    error = "Missing or multiple species tags within an event"
    )
```

Missing or multiple group count tags for a non-empty event:
```{r}
error_groupcount <- OSM_2024_data %>%
  
  group_by(site, episode_id, datasource) %>% # retrieve every image in an event
  
  filter(any(empty==FALSE)) %>% ## Select non-empty events
  
  # are there multiple for any of the group metrics, including NAs?
  filter( 
    n_distinct(group_count)>1 | 
    n_distinct(g_male)>1 |
    n_distinct(g_female)>1 |
    n_distinct(g_unknownsex)>1 |
    n_distinct(g_adult)>1 |
    n_distinct(g_yly)>1 |
    n_distinct(g_yoy)>1 |
    n_distinct(gunknownage)>1
    ) %>%
  
  mutate(
    error = "Missing or multiple group counts within a single event"
    )
```

Image is marked as not empty, but there are no frame counts:
```{r}
error_framecount_missing <- OSM_2024_data %>%
  
  # Non-empty, non-staff images
  filter(empty == FALSE,
         species != 'Staff') %>%
  
  filter(if_all(c(total, male, female, unknownsex, adult, yly, yoy, unknownage), is.na)) %>% # are all of them NA?
  
  mutate(
    error = "Image is not empty but has no frame counts... did you forget to tag it?"
    )
```

Any frame totals for a given image exceed group totals:
```{r}
error_framecount_high <- OSM_2024_data %>%
  
  group_by(site, episode_id, datasource) %>% # retrieve every image in an event
  
  # Are any of the frame counts higher than the corresponding group count?
  filter(
    total > group_count | 
    male > g_male |
    female > g_female |
    unknownsex > g_unknownsex |
    adult > g_adult |
    yly > g_yly |
    yoy > g_yoy |
    unknownage > gunknownage
    ) %>%
  
  mutate(
    error = "Frame totals exceed the group totals for the event"
    )
```

For species larger than coyote: sex totals, age totals, and animal totals should be equal.
```{r}
error_demographic <- OSM_2024_data %>%
  
  # big animals only. Can be expanded manually.
  filter(species %in% c('Moose', 'Black bear', 'Grey wolf', 'Caribou', 
                        'Mule deer', 'Grizzly bear', 'Unknown deer', 'Lynx', 
                        'Unknown ungulate', 'Unknown bear', 'Cougar' )
         ) %>%
  
  group_by(site, episode_id, datasource) %>% # retrieve every image in an event

  rowwise() %>% # Check each image individually. Helps the sum() function to work properly. 
  
  filter(
    # Calculate sums for frame sex columns and age columns and tests if they are equal. Two of these equalities would likely suffice, could drop the third. 
    sum(male, female, unknownsex, na.rm = TRUE) != sum(adult, yly, yoy, unknownage, na.rm = TRUE) |
    sum(male, female, unknownsex, na.rm = TRUE) != total |
    sum(adult, yly, yoy, unknownage, na.rm = TRUE) != total |
    
    # Calculate sums for group (g_) columns and tests if they are equal. Two of these equalities would likely suffice, could drop the third. 
    sum(g_male, g_female, g_unknownsex, na.rm = TRUE) != sum(g_adult, g_yly, g_yoy, gunknownage, na.rm = TRUE) |
    sum(g_male, g_female, g_unknownsex, na.rm = TRUE) != group_count |
    sum(g_adult, g_yly, g_yoy, gunknownage, na.rm = TRUE) != group_count
    ) %>%
  
  mutate(
    error = "Total, age, and sex columns are not equal"
    )
```

Missing coat colour for applicable species:
```{r}
# Missing coat colour
error_coatcolor <- OSM_2024_data %>%

  filter(empty == FALSE) %>%
  
  filter(
    species %in% c('Black bear', 'Red fox', 'Grey wolf', 'Grizzly bear' ) # the only species of interest for coat color. 
    ) %>%
  
  filter(
    coatcolour %>% 
      is.na(.) # is coat color NA?
    ) %>%
  
  mutate(
    error = "Coat color is missing"
    )
```

Missing antler counts for ungulates (with appropriate date ranges):
```{r}
error_antlercount <- OSM_2024_data %>%
  
  mutate(month=month(datetime)) %>% # extract month as a helper column
  
  filter(
    empty == FALSE # non empty images
    ) %>%

  filter(
    (species == 'Moose' & male>=1 & month %in% 9:10) | # Male moose Sept-Oct
    (species == 'Caribou' & month %in% 8:10) | # Any sex caribou Aug-Oct
    (species == 'White-tailed deer' & male>=1 & month %in% 8:12) | # Male deer Aug-Dec
    (species == 'Mule deer' & male>=1 & month %in% 8:12) # Male deer Aug-Dec
    ) %>%
  
  filter(
    is.na(leftantler) | # Missing values for any of these. Even if count is "unknown", minimum should be selected. 
    is.na(rightantler) | 
    is.na(lcount) | 
    is.na(rcount)
    ) %>%
  
  mutate(
    error = "Antler count is missing"
    )
```

Entirely empty events with a species tag (either the 'empty' or the species is wrong):
```{r}
error_emptyevent <- OSM_2024_data %>%
  
  group_by(site, episode_id, datasource) %>% # retrieve every image in an event
  
  filter(all(empty == TRUE), # Are all images tagged as empty?
         any(!is.na(species)), # Do any images have a species tag?
         species != 'Staff' # Non-staff events only
    ) %>%
  
  mutate(
    error = "Every image in an event is classified as empty, but a species tag is present!"
    )
```

Did not get tagged as a species or as an empty image. Either 1) the image didn't get tagged, or 2) an entire animal event is missing the species tag.   
```{r}
error_missingspecies_untagged <- OSM_2024_data %>%
  
  group_by(site, episode_id, datasource) %>% # retrieve every image in an event
  
  filter(any(empty == FALSE), # at least one image tagged as non-empty
    is.na(species) # all species values are empty.
    ) %>%

  mutate(
    error = "Forgot to enter the species tag, OR this image didn't get classified."
    )
```

Combine all the errors into a single spreadsheet and export:
```{r}
# Export the errors to a spreadsheet. As per the other sections. 
OSM_2024_error_image <- bind_rows(error_speciestag, 
                                  error_groupcount, 
                                  error_framecount_high,
                                  error_framecount_missing, 
                                  error_demographic, 
                                  error_coatcolor, 
                                  error_antlercount,
                                  error_emptyevent,
                                  error_missingspecies_untagged) %>%
  ungroup() %>%
  select(site, file, datetime, episode_id, classifier, species, error, comments, fullpath)


if(nrow(OSM_2024_error_image) == 0){
  writeLines("Lucky you! No image errors in the dataset.")
} else {
writexl::write_xlsx(OSM_2024_error_image, "./OSM_2024_error_image.xlsx", format_headers = TRUE)
writeLines("Image errors are likely present. Please refer to the excel sheet and correct in Timelapse directly.")
}

rm(error_speciestag, error_groupcount, error_framecount_high, error_framecount_missing, error_demographic, error_coatcolor, error_antlercount, error_emptyevent, error_missingspecies_untagged) # clean up
```


