---
title: "ACME Wildlife Detection and Covariate Processing"
author: "Marissa A. Dyck and Aidan M. Brushett"
date: "2025-02-09"
output:
  html_document:
    theme: default
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
```

# Before you begin

## Notes

A few notes about this script.

If you are running this with the 2024-2025 data make sure you download the whole (OSM_2024-2025 GitHub repository)[https://github.com/ACMElabUvic/OSM_2024-2025] from the ACMElabUvic GitHub. This will ensure you have all the files, data, and proper folder structure you will need to run this code and associated analyses.

Also make sure you open RStudio through the R project (OSM_2024-2025.Rproj) this will automatically set your working directory to the correct place (wherever you saved the repository) and ensure you don't have to change the file paths for some of the data. 

If you have question please email the most recent author, currently   

Marissa A. Dyck   
Postdoctoral research fellow    
University of Victoria    
School of Environmental Studies     
Email: [marissadyck17@gmail.com](marissadyck17@gmail.com)     

**and**

Aidan M. Brushett  
M.Sc. student
University of Victoria    
School of Environmental Studies     
Email: [aidanbrushett@uvic.ca](aidanbrushett@uvic.ca)     

## Netdrive access

This script relies on the user having access to the ACME lab Netdrive (you can view the .html output of this file if you don't have access and just want to see what the script did).

Helpful instructions for connecting to and navigating the Netdrive can also be found here: [https://docs.google.com/document/d/1Z72IrlIXO8MUHCoVztcMrMdL10R2tHrHThEgfow1Cu0/edit ](https://docs.google.com/document/d/1Z72IrlIXO8MUHCoVztcMrMdL10R2tHrHThEgfow1Cu0/edit).


## R and RStudio

Before starting you should ensure you have the latest version of R and RStudio downloaded. This code was generated under R version 4.2.3 and with RStudio version 2024.04.2+764.    

You can download R and RStudio [HERE](https://posit.co/download/rstudio-desktop/)   


## R markdown

This script is written in R markdown and thus uses a mix of coding markup languages and R. If you are planning to run this script with new data or make any modifications you will want to be familiar with some basics of R markdown.

Below is an R markdown cheatsheet to help you get started,    
[R markdown cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)    

# Setup

## Install packages

If you don't already have the following packages installed, use the code below to install them. *NOTE this will not run automatically as eval=FALSE is included in the chunk setup (i.e. I don't want it to run every time I run this code since I have the packages installed)

```{r install packages, eval=FALSE}

install.packages('tidyverse') 
install.packages('withr') 
```

## Load libraries

Then load the packages to your library.

```{r libraries}
rm(list=ls()) # start with a fresh global environment
library('tidyverse') # data tidying, visualization, and much more; this will load all tidyverse packages, can see complete list using tidyverse_packages()
library('withr') # used to temporarily set wd
```

# 1. Import 2024 camera operability 

We can check the length each camera was operating using the cleaned deployment data, this is important for calculating the proportional presence/absences for analysis later on so we need to make sure nothing looks inaccurate here. This was done as part of the Timelapse error checking script, but it's not a bad idea to do it again. 

Let's plot the camera operability with `ggplot()` to look at this

```{r camera operability}

# if starting from this point read in data
deploy <- read_csv('data/raw/OSM_deployment_2024.csv') %>% 
  
  drop_na(site) %>%
  
  # make sure site re-reads in as a factor to compare with other data sets
  mutate(site = as.factor(site),
         end_date = as.Date(end_date))


str(deploy)


# create graph of camera operability

ggplot(deploy, aes(color = array)) +
  
  geom_segment(aes(x = start_date, 
                   xend = end_date,
                   y = site, 
                   yend = site)) +
  
  theme(axis.text = element_text(size = 6))

```

Looks good! Bye again lol...
```{r}
rm(deploy)
```


# 2. Import all years of site coordinates

We will get these from the site deployment datasheets, originally sourced from the NetDrive. 

```{r}
osm_coords <- list.files(path="./data/raw/", 
                               pattern="OSM_Deployment_Site_", 
                               full.names = TRUE) %>%
      map_dfr(.,
          ~read_csv(.x,
                   col_types = cols(Array = col_factor(),
                                    Site = col_factor(),
                                    Lat = col_number(),
                                    Long = col_number())
                   ) %>%
            
            set_names(names(.) %>% tolower()) %>%
            
            select(-array) %>%
            
            mutate(
              site = gsub("-", "_", site) %>% # First, replace the hyphen with underscore
                gsub("_0+", "_", .) %>% # Then, remove leading zeros after the underscore
                gsub(" ", "", .) %>% # Then, remove any spaces
                gsub("LU0", "LU", .) %>% # Remove leading zeros from LU names for consistency
                as.factor(.), # Then, convert back to factor
               ) %>%
            
            # Split only at the first underscore
            extract(site, into = c("array", "camera"), 
                    regex = "^([^_]+)_(.*)$",
                    remove = F) %>%
      
            # specify format of new columns
            mutate(
              array = as.factor(array),
              camera = as.factor(camera)
              ) %>%
            
            select(array, site, lat, long)
      )
```

Get UTM coordinates (NAD83 UTM 12N):
                     
```{r}
# assuming lat/long was extracted in NAD83
osm_coords <- st_as_sf(osm_coords, coords = c("long", "lat"), crs = 4269) %>% 
  
  mutate(long = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2]) %>%
  
  st_transform(osm_coords, crs = 26912) %>%
  
  mutate(easting_12n = st_coordinates(.)[,1],
         northing_12n = st_coordinates(.)[,2]) %>%
  
  st_drop_geometry(.)

st_crs(osm_coords)
```

# 3. Import all years of deployment data. 

Read in the deployment data for everything.  Useful to have this in one place. 

```{r}
osm_deploy_multi_year <- list.files(path="./data/raw/", 
                               pattern="OSM_deployment_20", 
                               full.names = TRUE) %>%
      map_dfr(.,
          ~read_csv(.x,
                    col_types = cols(end_date = col_date()))
                   ) %>%
  
      drop_na(site) %>%
  
      mutate(site = gsub("-", "_", site) %>% # First, replace the hyphen with underscore
               gsub("_0+", "_", .) %>% # Then, remove leading zeros after the underscore
               gsub(" ", "", .) %>% # Then, remove any spaces
               gsub("LU0", "LU", .) %>% # Remove leading zeros from LU names for consistency
               as.factor(.), # Then, convert back to factor
             ) %>%
  
      select(-array) %>%
  
      # Split only at the first underscore
      extract(site, into = c("array", "camera"), 
                regex = "^([^_]+)_(.*)$",
                remove = F) %>%
  
  # make sure site re-reads in as a factor to compare with other data sets
  mutate(site = as.factor(site),
         array = as.factor(array)) %>%
  
  select(-camera)
```

Append the deployment history and site coordinates into a single file and save it. 

```{r}
osm_coords_deploy <- osm_coords %>%
  
  left_join(osm_deploy_multi_year, by = c("site", "array"))

summary(osm_coords_deploy)
```
                     
Save the file: 

```{r}
write_csv(osm_coords_deploy, "./data/processed/OSM_site_coordinates_2021-2024.csv")

rm(osm_deploy_multi_year, osm_coords)
```


# 4. Import and clean all years of timelapse files

First, import the images:

``` {r}
osm_images <- list.files(path="./data/raw/",
                         pattern="OSM_timelapse_20",
                         full.names = TRUE) %>%
      map_dfr(.,
          ~read_csv(.x,
                    col_types = cols(datetime = col_datetime(),
                                     site = col_character(),
                                     total = col_integer(),
                                     group_count = col_integer(), 
                                     comments = col_character(),
                                     species = col_factor(),
                                     otherspecify = col_character(),
                                     snow = col_factor(),
                                     empty = col_logical(),
                                     camera = col_character(),
                                     leftantler = col_character(),
                                     rightantler = col_character()
                                     ))) %>%
  
            mutate(
              site = gsub("-", "_", site) %>% # First, replace the hyphen with underscore
                gsub("_0+", "_", .) %>% # Then, remove leading zeros after the underscore
                gsub(" ", "", .) %>% # Then, remove any spaces
                gsub("LU0", "LU", .) %>% # Remove leading zeros from LU names for consistency
                as.factor(.), # Then, convert back to factor
               ) %>% 
  
              # Split only at the first underscore
            extract(site, into = c("array", "camera"), 
                    regex = "^([^_]+)_(.*)$",
                    remove = F) %>%
      
            # specify format of new columns
            mutate(
              array = as.factor(array)
              ) %>%
  
        select(-camera)

str(osm_images)

```

Let's do some rough checks:

```{r}
unique(osm_images$species)
unique(osm_images$array)
unique(osm_images$site)
```

There are some problems that should be fixed (mule deer, arctic hare, elk, etc.). We can revisit this later...

The missing array value is only one row so we'll drop it. We'll also do a couple other quick and dirty fixes here:

```{r}
osm_images <- osm_images %>% 
  
  # Drop the bad row
  drop_na(array) %>%
  
  # Replace all NAs for species counts with 0s to make Boolean-based filtering easier
  mutate(
    across(total:gunknownage, ~replace_na(.x, 0))
  ) %>%
  
  mutate(species = case_when(
          species == "Arctic hare" ~ "Snowshoe hare", 
          species == "Mule deer" ~ "White-tailed deer", 
          TRUE ~ species
  ))
```

Let's do those same checks again:

```{r}
unique(osm_images$species)
unique(osm_images$array)
unique(osm_images$site)
```

Good enough for government work. Andrew's data paper code would be very valuable here to get this cleaned up!

For now we can export the imagery:

```{r}
write_csv(osm_images, "./data/processed/OSM_timelapse_2021-2024.csv")
```


# 5. Determine camera operability for all years:

This deployment data is confirmed as correct. It does NOT account for snow cover, which we will do manually below using the image data. 

```{r}
osm_operating <- osm_coords_deploy %>% 
  
  drop_na(start_date, end_date) %>%
  
  rowwise() %>%
  
  # create a list sequence of dates for each site from the first to the last day of operability
  mutate(date = list(seq(from = date(start_date), to = date(end_date), by="day"))) %>% 
    
  unnest(date) %>%
  
  select(array, site, date)
```

We also need to modify the OSM operating dataframe to account for days where the camera was fully covered by snow or otherwise obstructed (cases where `cameramalfunction == *'Fully obscured'*`). Let's identify those from the raw timelapse data and remove them from `osm_operating`.

```{r}
osm_obscured <- osm_images %>%
  
  filter(cameramalfunction == 'Fully obscured') %>%
  
  mutate(date = as.Date(datetime)) %>%
  
  select(array, site, date) %>% 
  
  distinct()
```

Let's remove those dates:

```{r}
osm_operating <- osm_operating %>%
  
  anti_join(osm_obscured) %>%
  
  mutate(year = year(date), 
         month = month(date))
                                 
# Save the operating days file
#write_csv(osm_operating, "./data/raw/OSM_operating_days.csv")
```

Write the file:

```{r}
write_csv(osm_operating, "./data/processed/OSM_operating_days_2021-2024.csv")
rm(osm_obscured)
```


# 6. Independent detections

For the models we want to run, we need data for the independent detections of each species at each site. We've defined independent detections as those at least 30 minutes apart. 

The data manipulation and loop below will do this for us.

```{r independent detections}

# we will continue working with the OSM_2023_data but you could start here and import the data again if needed but you'd have to change the structure of a couple variables that get re-read in wrong after the export and import process. Can uncomment and use the code below to do that.

# can select code chunk and use command + shift + c to uncomment or comment a large portion of code

# OSM_2022_data_fixed <- read_csv('data/processed/OSM_2022_timelapse.csv')
# 
# # ignoring parsing issues warning, this is just referring to some columns it's expecting logical data for and they contain characters.
# 
# # check internal structure, even though we specified everything above with the fixed data when we export and import R often reads the variables in wrong again
# str(OSM_2022_data_fixed) 
# 
# # datatime read correctly but we will need to change site and species to factors again
# 
# OSM_2022_data_fixed <- OSM_2022_data_fixed %>% 
#   
#   mutate(species = as.factor(species),
#          site = as.factor(site))
  

# set the independent detection threshold to 30 minutes
threshold <- 30 

# prep the data for calculating independent detections
detections <- osm_images %>% 
  
  # remove rows with no species info
  drop_na(species) %>% 
  
  # We want to avoid using empty group images where no animal is actually present. This could falsely alter independent detections. 
  filter(empty == FALSE) %>% 
  
  # select only variables of interest
  select(array,
         site,
         species,
         datetime,
         month,
         year) %>% 
  
  # now we need to create a new variable called timediff
  # first make sure data are arrange in proper order
  arrange(site, species, datetime) %>% # this will NOT work if not in correct order (early-late date)
  
  # create groups for each species at each site
  group_by(species, site) %>%
  
  # create new variable timediff that will calculate the difference 
  mutate(timediff = as.numeric(difftime(datetime,lag(datetime),
                                        units = "mins")),
        # Then, create an event_ID that groups detections and identifies when photos are >30min apart. 
        event_id = cumsum(if_else(is.na(timediff) | timediff > threshold,  # Create episode IDs to distinguish detections 
                                  1, 
                                  0))
        )

# now create a new data frame with a single row for each event. 
# The timestamp corresponds to the EARLIEST image in the event. Other code could be added 
osm_indet <- detections %>% 
  
  group_by(array, species, site, event_id, year, month) %>%
  
  summarize(datetime = min(datetime)) %>%
  
  mutate(species = tolower(species))

## This code can be subbed in to pull the first and last timestamps. Useful for calculating COARSE event duration. 
#  summarize(event_start = min(datetime), event_end = max(datetime))
```

## Save independent detections

Let's also save this data file for later

```{r save detection data}
write_csv(osm_indet,
          './data/processed/OSM_independent_detections_2021-2024.csv')

rm(detections)
```



# 7. Response variables


## 7.0. What species do we want to have??

```{r}
 focal_species <- c('black bear',
                    'caribou',
                    'cougar',
                    'coyote',
                    'fisher',
                    'grey wolf',
                    'lynx',
                    'moose',
                    'red fox',
                    'white-tailed deer',
                    'wolverine',
                    'snowshoe hare', 
                    'red squirrel'
                    )
```


## 7.1. Monthly independent detections and presence-absence

Bin the operating data monthly and retain months with \>15 days of data.

```{r}
operating_months <- osm_operating %>%
  
  group_by(array, site, month, year) %>%
  
  summarize(n_days = n()) %>%
  
  filter(n_days >= 15) # retain months where there are 15 days of data only.
```

Then, let's bin the independent detections data into months. This data frame will only contain the positive detections. i.e., if a species was undetected for a given site/month, there will be no row.

```{r}
indet_months <- osm_indet %>%
  
  filter(species %in% focal_species) %>%

  group_by(array, site, species, month, year) %>%
  
  summarize(presence = 1,
            detections = n())
```

Then, let's merge this to the operating data using left_join. This will ensure that our 'absence' cases now show up in the dataset.

```{r}
presence_months <- operating_months %>%
  
  # Ensure that we have a new row for each species x month combination. 
  crossing(species = focal_species) %>%
  
  left_join(indet_months, by = c("array", "site", "species", "month", "year")) %>%
  
  # Fill missing presence and detections with 0
  mutate(presence = replace_na(presence, 0),
         detections = replace_na(detections, 0),
         species = as.factor(species)) %>%
  
  select(-n_days)

summary(presence_months)
```

Save the file:

```{r}
write_csv(presence_months, "./data/processed/OSM_response_monthly_presence_detections_2021-2024.csv")
```

## 7.2. Monthly proportional presence-absence

We will now create a proportional monthly variable for each species by determining how many months an animal was present, divided by the total months of data available.

```{r, message=FALSE}
proportional_months <- presence_months %>%
  
  group_by(array, site, species) %>%
  
  summarize(months_active = n(),
            months_present = sum(presence),
            .groups = 'drop')
```

Let's do a couple quick checks to make sure this worked. First, a visual inspection to make sure we have the same number of rows for each species and a reasonable range of active dates. Then, we'll check if any of the presence values exceed the active values, which would indicate an error.

```{r}
summary(proportional_months)

proportional_months %>%
  filter(months_present > months_active)
```

Now let's pivot the data to wide format :) We will also rename the species since they're now becoming column names and spaces or hyphens aren't acceptable. 

```{r, message=FALSE}
proportional_months <- proportional_months %>%
  
  # pivot the data wider so there is a column for each species and 1 row per site
  pivot_wider(names_from = species,
              values_from = months_present) %>% 
  
  # replace NAs with zeros in all species columns
  mutate(across(
    where(is.numeric),
    ~ replace_na(., 0))) %>%
  
  # ensure all species columns are numeric not integer
  mutate_if(is.integer,
            as.numeric) %>%
  
  # set the column names to lower case and replace the spaces with '_' (these are both personal preferences of mine)
  set_names(
    names(.) %>% 
      tolower()%>% 
      str_replace_all(pattern = ' ',
                      replacement = '_') %>%
      str_replace_all(pattern = '-',
                      replacement = "_"))
```

Now we can run the function below to create a second column for each species that will represent the number of absences (months the species was not detected) at each camera from the active months. This function was adopted from Marissa Dyck's OSM 2022-2023 image processing script, found on the ACME GitHub.

```{r}
# replace all non-letter digits with underscores to match the column names in the response variable. 
focal_species_colnames <- focal_species %>% 
  str_replace_all(., pattern = ' ', replacement = '_') %>%
  str_replace_all(., pattern = '-', replacement = '_')
  
# first convert data to data frame not a tibble for function to work
proportional_months <- as.data.frame(proportional_months)

# create a vector of the species columns for the loop
# update our focal species list to match the new column names
for (species in focal_species_colnames) {
  
  if (is.numeric(proportional_months[,species]) & is.numeric(proportional_months$months_active)) {
    col_absence <- paste0("absent_", species)
    
    proportional_months[col_absence] <- proportional_months$months_active - proportional_months[,species]

  }
}

str(proportional_months)
```

Let's adjust the black bear data to accommodate hibernation. We want bear data to represent only April to November.

```{r}
operating_months_bears <- operating_months %>%
  
  filter(month %in% c("4", "5", "6", "7", "8", "9", "10", "11")) %>%
  
  group_by(array, site) %>%
  
  summarize(months_active_bears = n(),
            .groups = 'drop')

# Update the months_active field for bears
proportional_months <- proportional_months %>%
  
  left_join(operating_months_bears, by = c('array', 'site')) %>%
  
  # overwrite absent black bear column
  mutate(absent_black_bear = months_active_bears - black_bear) %>%
  
  # get rid of unnecessary columns for active months
  select(-months_active_bears)

```

Save the file:

```{r}
write_csv(proportional_months, "./data/processed/OSM_response_monthly_proportional_presence_2021-2024.csv")
```

Clean up our workspace:

```{r}
rm(operating_months_bears, operating_months, indet_months)
```

## 7.3. Weekly independent detections and presence-absence

Bin the operating data weekly and retain weeks with \>4 days of data. There is a handy function to do this!! `floor_date()` takes a date-time object and rounds it down to the nearest boundary of the specified time unit, starting on Sunday (default case). In this step, we will round down our operating data to the nearest week so we can bin it. We'll re-extract the month and year fields later.

```{r}
operating_weeks <- osm_operating %>%
  
  mutate(week = floor_date(date, unit = "week")) %>%
  
  group_by(array, site, week) %>%
  
  summarize(n_days = n()) %>%
  
  filter(n_days >= 4) # retain weeks where there are 15 days of data only.
```

Then, let's bin the independent detections data into weeks. This data frame will only contain the positive detections. i.e., if a species was undetected for a given site/week, there will be no row.

```{r}
indet_weeks <- osm_indet %>%
  
  filter(species %in% focal_species) %>%

  mutate(week = floor_date(datetime, unit = "week")) %>%
  
  group_by(array, site, week, species) %>%
  
  summarize(presence = 1,
            detections = n(),
            .groups = 'drop')
```

Then, let's merge this to the operating data using left_join. This will ensure that our 'absence' cases now show up in the dataset.

```{r}
presence_weeks <- operating_weeks %>%
  
  # Ensure that we have a new row for each species x week combination. 
  crossing(species = focal_species) %>%
  
  left_join(indet_weeks, by = c("array", "site", "species", "week")) %>%
  
  # Fill missing presence and detections with 0
  mutate(presence = replace_na(presence, 0),
         detections = replace_na(detections, 0),
         species = as.factor(species)) %>%
  
  mutate(month = month(week),
         year = year(week)) %>%
  
  select(-n_days)

summary(presence_weeks)
```

Save the file:

```{r}
write_csv(presence_weeks, "./data/processed/OSM_response_weekly_presence_detections_2021-2024.csv")
```

## 7.4. Weekly proportional presence-absence

We will now create a proportional weekly variable for each species by determining how many weeks an animal was present, divided by the total weeks of data available.

```{r, message=FALSE}
proportional_weeks <- presence_weeks %>%
  
  group_by(array, site, species) %>%
  
  summarize(weeks_active = n(),
            weeks_present = sum(presence),
            .groups = 'drop')
```

Let's do a couple quick checks to make sure this worked. First, a visual inspection to make sure we have the same number of rows for each species and a reasonable range of active dates. Then, we'll check if any of the presence values exceed the active values, which would indicate an error.

```{r}
summary(proportional_weeks)

proportional_weeks %>%
  filter(weeks_present > weeks_active)
```

Now let's pivot the data to wide format :)

```{r, message=FALSE}
proportional_weeks <- proportional_weeks %>%
  
  # pivot the data wider so there is a column for each species and 1 row per site
  pivot_wider(names_from = species,
              values_from = weeks_present) %>% 
  
  # replace NAs with zeros in all species columns
  mutate(across(
    where(is.numeric),
    ~ replace_na(., 0))) %>%
  
  # ensure all species columns are numeric not integer
  mutate_if(is.integer,
            as.numeric) %>%
  
  # set the column names to lower case and replace the spaces with '_' (these are both personal preferences of mine)
  set_names(
    names(.) %>% 
      tolower()%>% 
      str_replace_all(pattern = ' ',
                      replacement = '_') %>%
      str_replace_all(pattern = '-',
                      replacement = "_"))
```

Now we can run the function below to create a second column for each species that will represent the number of absences (weeks the species was not detected) at each camera from the active weeks. This function was adopted from Marissa Dyck's OSM 2022-2023 image processing script, found on the ACME GitHub.

```{r}
# first convert data to data frame not a tibble for function to work
proportional_weeks <- as.data.frame(proportional_weeks)

# create a vector of the species columns for the loop
# update our focal species list to match the new column names
for (species in focal_species_colnames) {
  
  if (is.numeric(proportional_weeks[,species]) & is.numeric(proportional_weeks$weeks_active)) {
    col_absence <- paste0("absent_", species)
    
    proportional_weeks[col_absence] <- proportional_weeks$weeks_active - proportional_weeks[,species]

  }
}

str(proportional_weeks)
```

Let's adjust the black bear data to accommodate hibernation. We want bear data to represent only April to November.

```{r}
operating_weeks_bears <- operating_weeks %>% 
  
  mutate(month = month(week)) %>%
  
  filter(month %in% c(4:11)) %>%
  
  group_by(array, site) %>%
  
  summarize(weeks_active_bears = n(),
            .groups = 'drop')

# Update the weeks_active field for bears
proportional_weeks <- proportional_weeks %>%
  
  left_join(operating_weeks_bears, by = c('array', 'site')) %>%
  
  # overwrite absent black bear column
  mutate(absent_black_bear = weeks_active_bears - black_bear) %>%
  
  # get rid of unnecessary columns for active weeks
  select(-weeks_active_bears)

```

Save the file:

```{r}
write_csv(proportional_weeks, "./data/processed/OSM_response_weekly_proportional_presence_2021-2024.csv")
```

Clean up our workspace:

```{r}
rm(operating_weeks_bears, operating_weeks, indet_weeks)
```

## 7.5. Total independent detections controlled for with survey effort

```{r fig.height=8, fig.width=10}
total_indet <- osm_indet %>% 
  
  filter(species %in% focal_species) %>%
  
  mutate(date = date(datetime)) %>%
  
  semi_join(., osm_operating, by = c("site", "date")) %>%
  
  group_by(array, site, species) %>%
  
  summarize(detections = n())


total_detections <- osm_operating %>%
  
  group_by(array, site) %>%
  
  summarize(camera_days = n()) %>%
  
  crossing(species = focal_species) %>%
  
  left_join(total_indet, by = c("array", "site", "species")) %>%
  
  mutate(detections = replace_na(detections, 0),
         presence = ifelse(detections > 0, 1, 0))
```

Save the file:

```{r}
write_csv(total_detections, "./data/processed/OSM_response_total_detections_per_effort_2021-2024.csv")
```



End script








