---
title: "Extract landcover covariates"
author: "Aidan Brushett"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: default
    toc: yes
    toc_float: yes
---

```{r setup, include=TRUE}
# This is set to false just for quick knitting
knitr::opts_chunk$set(echo = T,
                      eval = T)
```

```{r}
library(sf) # for points and buffering
library(tidyverse)
rm(list = ls())
```

# 1. Read in the raw site coordinates:

I'll read in all the sites into a single file first: 

```{r}
sites_osm <- read_csv("./data/processed/OSM_site_coordinates_2021-2024.csv") %>%
  
  st_as_sf(., coords = c("easting_12n", "northing_12n"), crs = 26912) %>%
  
  mutate(year = year(start_date))
```

We will also subset based on survey year, since 2021 and 2022 use separate human footprint databases from ABMI. This splits all of them by year of deployment date!

```{r}
# Sampled 2022 onwards
sites_osm_2022_2023_2024 <- sites_osm %>% 
  
  filter(year >= 2022)


# Sampled 2021 
sites_osm_2021 <- sites_osm %>% 
  
  filter(year == 2021)

```



# 2. Forest inventory data

## 2.1. Load, clip the forest inventory data

We can do this for all sites at once (i.e. not split by survey year, since it's the same dataset for all arrays). First we'll buffer the camera sites so we can tell what 

```{r}
#For combining multiple sf objects into 1 (not needed currently other than for renaming to `sites`)
sites <- bind_rows(sites_osm) #, sites_christina)

# Check it out on a map
plot(sites)

# buffer the sites so that we can filter the landcover dataset and reduce file size
# An 11km buffer will encompass all features for the buffer radii that we are extracting. 
sites_buffered <- st_buffer(sites, dist = 11000) %>%
  st_union()

# check it out on a map
plot(sites_buffered[1])
```

Now we can load in and clip the geodatabase containing the 2020 SBFI. This takes quite a while depending on file size. This is NOT in the Github, but can be stored on individual users' computers from the public download based on the costs. 

```{r}
# load landcover layer (will be slow)
landcover_unclipped <- st_read("C:/Users/aidanbrushett/Desktop/projects/OSM_covariate_extraction/SBFI2020_boreal_plains.gdb", layer = "SBFI_2020_polygons")

landcover <- landcover_unclipped %>%
  # Reproject to UTM 11N, same as the sites
  st_transform(., crs = 26912) %>%
  # Clip to our camera buffers
  filter(lengths(st_intersects(., sites_buffered)) > 0)

#make sure it's all projected correctly 
st_crs(landcover, describe = TRUE)
st_crs(sites, describe = TRUE)
plot(landcover[1])

# clean up environment
rm(landcover_unclipped, sites_buffered)
gc()

# double check that they are all format "sf" and "dataframe"
str(sites)
str(landcover)
```

## 2.2. Reshape data for key variables 

First, tree species composition. This is a total pain because it's currently stored based on ranked columns rather than columns for each species name. We are going to take the value from each Species 1, Species *n* column, etc. and sum them together, then transform the data so we get a single percentage value for each polygon (row) for each tree species (column)


Get a list of the tree species present in the dataset in vector format. 

```{r}
tree_spp <- landcover %>%
  
  # Faster run time
  st_drop_geometry() %>%
  
  # These are the only ones we want
  select(SPECIES_1,
         SPECIES_2,
         SPECIES_3,
         SPECIES_4,
         SPECIES_5) %>%
  
  unlist(.) %>%   # In a big ugly vector
  
  unique(.) %>% # What are the unique tree species present?

  .[ . != " "] # Remove empty strings
```

Re-pivot the tree data for each polygon so that the species names are now the column names. This can take a while and is a bit clunky. 

```{r}
tree_percent <- landcover %>%
  
  st_drop_geometry() %>%
  
  # Select the species columns that have the name and % comp only
  select(ID,
         contains("SPECIES") & 
           !contains("_CML"),
         -SPECIES_NUMBER,
         -SPECIES_CONIFEROUS_PERC
         ) %>%
  # Convert all columns to character
  mutate(across(everything(), as.character)) %>%  
  
  # The start of some funky data reshaping
  pivot_longer(cols = starts_with("SPECIES_"),
               names_to = "temp",
               values_to = "val") %>%
  
  # What rank is the species? We don't actually care but this helps keep things tidy/avoids duplicates when we pivot wider later
  mutate(spp_rank = as.numeric(str_extract(temp, "\\d+"))) %>%

  # Remove the numbers so that all the former rank columns are just called "species"
  mutate(temp = str_replace_all(temp, "_\\d+", ""),
         temp = str_replace_all(temp, "SPECIES_PERC", "PERCENTAGE")) %>%
  
  # Pivot wider so we now have a species column and a percent column
  pivot_wider(names_from = temp, values_from = val) %>%
  
  # Only want the non-zero percents
  filter(PERCENTAGE > 0) %>%

  # Pivot wider again to get our species and comp values!!
  pivot_wider(names_from = SPECIES, values_from = PERCENTAGE, values_fill = "") %>%
  
  select(-spp_rank) %>%
  
  # Convert back to numeric
  mutate(across(everything(), as.numeric)) %>%
  
  # Replace NAs
  mutate(across(everything(), ~ replace(., is.na(.), 0))) %>%
  
  # Group by polygon again
  group_by(ID) %>%

  # Take the sum of the % comp for all the species
  summarize(across(everything(), sum))
```

Now we can repeat a similar process for timber harvest, extracted by year. 

```{r}
# Get harvest percent
harvest_percent <- landcover %>% 
  
  # Runs faster without coordinates
  st_drop_geometry(.) %>%
  
  # Year and amount columns, plus ID for pivoting/joining later
  select(ID,
         contains("DISTURB_HARVEST")) %>%
  
  # Reshape to a column for each year
  pivot_wider(names_from = DISTURB_HARVEST_YEAR, 
              values_from = DISTURB_HARVEST_PERC, 
              values_fill = 0) %>%
  
  select(-`0`) %>%
  
  # clean up the names
  rename_with(., ~paste0("HARVEST_PCT_", .), .cols = !contains("ID")) %>%
  
  # order by year
  select(ID, sort(names(.)))
```

Now we can repeat a similar process for percent fire, extracted by year. 

```{r}
# Get fire percent
fire_percent <- landcover %>% 
  
  st_drop_geometry(.) %>%
  
  # Year and proportion columns, plus ID for pivoting/joining later
  select(ID,
         DISTURB_FIRE_YEAR,
         DISTURB_FIRE_PERC) %>%
  
  # Reshape to a column for each year
  pivot_wider(names_from = DISTURB_FIRE_YEAR, 
              values_from = DISTURB_FIRE_PERC, 
              values_fill = 0) %>%
  
  select(-`0`) %>%

  # clean up the names
  rename_with(., ~paste0("FIRE_PCT_", .), .cols = !contains("ID")) %>%
    
  select(ID, sort(names(.)))
```

Intensity of fire for a given polygon. Same code as above just with a new variable of interest. 

```{r}
fire_intensity <- landcover %>% 
  
  st_drop_geometry(.) %>%
  
  select(ID,
         DISTURB_FIRE_MAGNITUDE_AVG,
         DISTURB_FIRE_YEAR) %>%
  
  pivot_wider(names_from = DISTURB_FIRE_YEAR, 
              values_from = DISTURB_FIRE_MAGNITUDE_AVG, 
              values_fill = 0) %>%
  
  select(-`0`) %>%
  
  rename_with(., ~paste0("FIRE_MAG_", .), .cols = !contains("ID")) %>%
    
  select(ID, sort(names(.)))
```

Left join all of the new covariates to the original spatial landcover data and replace NAs induced by joining rows that contained no fire/harvest.

```{r}
landcover_recalculated <- landcover %>%
  
  left_join(tree_percent, by = "ID") %>%
  
  left_join(fire_percent, by = "ID") %>%
  
  left_join(fire_intensity, by = "ID") %>%
  
  left_join(harvest_percent, by = "ID") %>%
  
  # Replace NAs
  mutate(across(
        .cols = c(all_of(tree_spp), contains("FIRE_PCT"), contains("HARVEST_PCT")), 
        ~ replace(., is.na(.), 0))) %>%
  
  # Original tree species names were ugly, should fix later. 
  rename_with(., ~paste0(., "_PCT_OF_TREED"), .cols = all_of(tree_spp))

rm(tree_spp, tree_percent, fire_percent, fire_intensity, harvest_percent)
```

## 2.3. Identify variables to extract

This is mostly to clean it up since this is a pretty unwieldly data frame. Helps for debugging to only see the things of interest :)

```{r, message=FALSE, warning=FALSE}
landcover_metrics <- landcover_recalculated %>% 
  
  # Select the metrics that you want to calculate
  select(., 
         
         # Forest age
         contains("AGE_"),
         -AGE_AVG,
         -AGE_MIN,
         -AGE_MAX,
         -AGE_MEDIAN,
         -AGE_SD,
         
         # Landcover
         contains("LC_"),
         -LC_FAO_FOREST, 
         -LC_TREED, 
         -LC_WETLAND_VEGETATION,
         
         # Tree composition
         contains("_PCT_OF_TREED"),
         
         # Fire and harvest percent
         contains("FIRE_PCT"),
         contains("HARVEST_PCT")
         )
```

## 2.4. Extract data for the sites of interest

This sums the % cover within each polygon and is weighted by their proportion of area within the buffer.

-   e.g. a polygon takes up 0.50 of the buffer and was 50% conifer = 25% conifer total
-   an additional polygon takes up 0.25 of the buffer and was 80% conifer = 20% conifer total
-   the proportion cover for the total buffer around the site is 45% conifer. 
-   *within a polygon the vegetation attributes are assumed to be constant*

```{r, message=FALSE, warning=FALSE}
sites_landcover_metrics <- list() # To store results

# For progress bar
start_time <- Sys.time()

# A `for` loop is nicer for progress statements IMO vs. purrr...
for(i in 1:nrow(sites)) {
  
  progress <- (i - 1) / nrow(sites)
  
  # Site we are extracting
  site <- sites[i,]

  # Progress bar
  cat("\r Extracting site ", i, " of ", nrow(sites), " (", site$site, "): ", 
      round(as.numeric(difftime(Sys.time(), start_time, units = "hours")) * (1 - progress) / progress, 2), 
      " hours remaining        ", sep = "")
  
  # Store results in a list
  sites_landcover_metrics[[i]] <- 
    
    # Buffer distances for which we are extracting using purrr
    purrr::map_dfr(c(50, 100, 200, 300, 400, seq(250, 5000, 250), 10000), ~
      
      # Clip the landcover to the site buffer               
      st_intersection(landcover_metrics,
                      st_buffer(site, .x)) %>%
      
      mutate(feature_area = st_area(.) %>% # Area of the feature within the buffer
               as.numeric(.),
             
             # Total area sampled. For the LANDCOVER ONLY DATA, we will remove NA areas from our 'total' area calculation. 
             # Obviously this doesn't make sense for HFI data.
             total_area = sum(feature_area, na.rm = TRUE) %>% 
               as.numeric(.),
             
             # Percent cover by a given feature
             percent_area = feature_area / total_area,
             
             # Values for 'proportion cover' adjusted by the amount of buffer area (see above Rmd notes)
             across(where(is.numeric) & !contains("_area"), ~ . / 100 * percent_area)) %>%
    
      # Runs faster without geometry
      st_drop_geometry(.) %>%
    
      # Sum across all features
      summarise(across(where(is.numeric) & !contains("_area"), sum, na.rm = TRUE)) %>%
    
      # Keep track of where the numbers came from!
      mutate(buffer_dist = .x,
             array = site$array,
             site = site$site)
      
    )
  
  # Clear some memory
  rm(site)
  gc()

}

# Save a backup in the same chunk
#save(sites_landcover_metrics, file = "./data/raw/OSM_SBFI2020_site_covariates.RData")
```

## 2.5. Save results

We'll save to the "./data/raw" folder even though this is technically a 'processed' output, since it will undergo lots more formatting later. 

```{r, message=FALSE, warning=FALSE}
sites_landcover_metrics_df <- bind_rows(sites_landcover_metrics) %>%

  relocate(array, site, buffer_dist)

write_csv(sites_landcover_metrics_df, "./data/raw/OSM_SBFI2020_site_covariates.csv")
```


***






# 3. Human footprint data 2021 -- for OSM sites in 2021 ONLY

Future years are extracted using the 2022 HFI for now. 

## 3.1. Import data

```{r}
# Camera locations (point)
sites <- bind_rows(sites_osm_2021)

# buffer sites so that we can filter the landcover data and reduce file size
sites_buffered <- st_buffer(sites, dist = 11000) %>%
  
  st_union()

plot(sites_buffered[1])

# Human footprint index:
# load human features layer (will be slow)
#st_layers("./maps/HFI2021_clipped.gdb")

hfi <- st_read("C:/Users/aidanbrushett/Desktop/projects/OSM_covariate_extraction/HFI2021_boreal_plains/HFI_2021_clipped.shp") %>%
  
  st_transform(., crs = 26912)


#make sure it's all projected correctly 
st_crs(hfi, describe = TRUE)
st_crs(sites, describe = TRUE)
#plot(hfi[1])

# clean up
rm(sites_buffered)
gc()

# double check that they are all format "sf" and "dataframe"
str(sites)
str(hfi)
```

## 3.2. Extract variables

```{r message=FALSE, warning=FALSE}
# Extract HFI (Human Footprint Inventory) covariates using site buffers
hfi_metrics <- hfi %>%
  select(OBJECTID, FEATURE_TY, YEAR) # Keep only relevant columns

sites_hfi_metrics <- list() # List to store output

start_time <- Sys.time() # Start timing the process

# Loop over each site in the 'sites' object
for(i in 1:nrow(sites)) {
  
  progress <- (i - 1) / nrow(sites) # Track progress
  
  site <- sites[i, ] # Extract the current site
  
  # Print progress to console with estimated time remaining
  cat("\r Extracting site ", i, " of ", nrow(sites), " (", site$site, "): ", 
      round(as.numeric(difftime(Sys.time(), start_time, units = "hours")) * (1 - progress) / progress, 2), 
      " hours remaining        ", sep = "")
  
  # Extract metrics for the current site across buffer distances
  sites_hfi_metrics[[i]] <-
    
    purrr::map_dfr(c(50, 100, 200, 300, 400, seq(250, 5000, 250), 10000), ~{
      
      site_buffer <- st_buffer(site, .x) # Create buffer around site
      
      st_intersection(hfi_metrics, site_buffer) %>% # Intersect with HFI layer
        
        # Calculate areas and percent cover
        mutate(feature_area = st_area(.) %>%
                 as.numeric(.), # Area of each intersected HFI polygon
               
               total_area = st_area(site_buffer) %>%
                 as.numeric(.), # Total area of buffer (unlike landcover, we use full area)
               
               percent_area = feature_area / total_area) %>% # Proportion of buffer occupied
        
        st_drop_geometry(.) %>% # Drop geometry for efficiency
        
        # Differentiate harvested areas by year
        mutate(FEATURE_TY = ifelse(str_detect(FEATURE_TY, "HARVEST-AREA"),
                                   paste0(FEATURE_TY, "-", YEAR),
                                   FEATURE_TY)) %>%
        
        # Group by array/site/feature type to summarize percent cover
        group_by(array, site, FEATURE_TY) %>%
        
        summarize(PCT_COVER = sum(percent_area),
                  buffer_dist = .x) %>%
        
        # Pivot to wide format: one row per site-buffer combo, features as columns
        pivot_wider(names_from = FEATURE_TY,
                    values_from = PCT_COVER,
                    values_fill = 0)
    })
  
  # Clean up memory after each iteration
  rm(site)
  gc()
}

#save(sites_hfi_metrics, file = "./data/output/OSM_HFI2021_site_covariates.RData")
```


## 3.3. Save results

We'll save to the "./data/raw" folder even though this is technically a 'processed' output, since it will undergo lots more formatting later. 

```{r message=FALSE}
# Some extra tidying. We need to deal with missing buffers/sites when there were no human features. 
sites_hfi_metrics_df <- bind_rows(sites_hfi_metrics) %>%
  
  ungroup() %>%

  relocate(array, site, buffer_dist, sort(names(.))) %>%
  
  select(-array) %>%
  
  complete(site, buffer_dist) %>%
  
  # Create some extra columns
  separate_wider_delim(site,
                       delim = '_',
                       names = c('array', 'camera'),
                       cols_remove = FALSE) %>%
  
  select(-camera) %>%
  
  # Replace NAs introduced by "complete"
  mutate(across(4:last_col(), ~replace(., is.na(.), 0)))

write_csv(sites_hfi_metrics_df, "./data/raw/OSM_HFI2021_site_covariates.csv")
```

***





# 4. Human footprint data 2022 -- for OSM sites in 2022 ONWARDS

Previous years are extracted using the 2021 HFI for now. 

## 4.1. Import data

```{r}
# Camera locations (point)
sites <- sites_osm_2022_2023_2024
# buffer sites so that we can filter the landcover data and reduce file size
sites_buffered <- st_buffer(sites, dist = 11000) %>%
  
  st_union()

plot(sites_buffered[1])

# Human footprint index:
# load human features layer (will be slow)
#st_layers("./maps/HFI2021_clipped.gdb")

hfi <- st_read("C:/Users/aidanbrushett/Desktop/projects/OSM_covariate_extraction/HFI2022_boreal_plains/HFI_2022_clipped.shp") %>%
  
  st_transform(., crs = 26912)


#make sure it's all projected correctly 
st_crs(hfi, describe = TRUE)
st_crs(sites, describe = TRUE)
#plot(hfi[1])

# clean up
rm(sites_buffered)
gc()

# double check that they are all format "sf" and "dataframe"
str(sites)
str(hfi)
```
Align the timber harvest names so that they correspond to the harvest names in the 2021 HFI. I don't know why they changed this...

```{r}
sort(unique(hfi$FEATURE_TY))

hfi <- hfi %>%
  
  mutate(FEATURE_TY = case_when(
    FEATURE_TY == "TIMBER-HARVEST-GREEN-AREA" ~ "HARVEST-AREA",
    FEATURE_TY == "TIMBER-HARVEST-WHITE-AREA" ~ "HARVEST-AREA-WHITE-ZONE",
    TRUE ~ FEATURE_TY))

sort(unique(hfi$FEATURE_TY))

```


## 4.2. Extract variables

```{r message=FALSE, warning=FALSE}
# Extract HFI (Human Footprint Inventory) covariates using site buffers
hfi_metrics <- hfi %>%
  select(OBJECTID, FEATURE_TY, YEAR) # Keep only relevant columns

sites_hfi_metrics <- list() # List to store output

start_time <- Sys.time() # Start timing the process

# Loop over each site in the 'sites' object
for(i in 1:nrow(sites)) {
  
  progress <- (i - 1) / nrow(sites) # Track progress
  
  site <- sites[i, ] # Extract the current site
  
  # Print progress to console with estimated time remaining
  cat("\r Extracting site ", i, " of ", nrow(sites), " (", site$site, "): ", 
      round(as.numeric(difftime(Sys.time(), start_time, units = "hours")) * (1 - progress) / progress, 2), 
      " hours remaining        ", sep = "")
  
  # Extract metrics for the current site across buffer distances
  sites_hfi_metrics[[i]] <-
    
    purrr::map_dfr(c(50, 100, 200, 300, 400, seq(250, 5000, 250), 10000), ~{
      
      site_buffer <- st_buffer(site, .x) # Create buffer around site
      
      st_intersection(hfi_metrics, site_buffer) %>% # Intersect with HFI layer
        
        # Calculate areas and percent cover
        mutate(feature_area = st_area(.) %>%
                 as.numeric(.), # Area of each intersected HFI polygon
               
               total_area = st_area(site_buffer) %>%
                 as.numeric(.), # Total area of buffer (unlike landcover, we use full area)
               
               percent_area = feature_area / total_area) %>% # Proportion of buffer occupied
        
        st_drop_geometry(.) %>% # Drop geometry for efficiency
        
        # Differentiate harvested areas by year
        mutate(FEATURE_TY = ifelse(str_detect(FEATURE_TY, "HARVEST-AREA"),
                                   paste0(FEATURE_TY, "-", YEAR),
                                   FEATURE_TY)) %>%
        
        # Group by array/site/feature type to summarize percent cover
        group_by(array, site, FEATURE_TY) %>%
        
        summarize(PCT_COVER = sum(percent_area),
                  buffer_dist = .x) %>%
        
        # Pivot to wide format: one row per site-buffer combo, features as columns
        pivot_wider(names_from = FEATURE_TY,
                    values_from = PCT_COVER,
                    values_fill = 0)
    })
  
  # Clean up memory after each iteration
  rm(site)
  gc()
}

#save(sites_hfi_metrics, file = "./data/output/OSM_HFI2022_site_covariates.RData")
```


## 4.3. Save results

We'll save to the "./data/raw" folder even though this is technically a 'processed' output, since it will undergo lots more formatting later. 

```{r message=FALSE}
# Some extra tidying. We need to deal with missing buffers/sites when there were no human features. 
sites_hfi_metrics_df <- bind_rows(sites_hfi_metrics) %>%
  
  ungroup() %>%

  relocate(array, site, buffer_dist, sort(names(.))) %>%
  
  select(-array) %>%
  
  complete(site, buffer_dist) %>%
  
  # Split only at the first underscore
  extract(site, into = c("array", "camera"), 
          regex = "^([^_]+)_(.*)$",
          remove = F) %>%
  
  select(-camera) %>%
  
  # Replace NAs introduced by "complete"
  mutate(across(4:last_col(), ~replace(., is.na(.), 0))) 


write_csv(sites_hfi_metrics_df, "./data/raw/OSM_HFI2022_site_covariates.csv")
```

***
