---
title: "Format covariates"
author: "Aidan Brushett"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: default
    toc: yes
    toc_float: yes
---

# Before you begin

Aidan Brushett
Ph.D. Student
University of Victoria    
School of Environmental Studies     
Email: [aidanbrushett@uvic.ca](aidanbrushett@uvic.ca)

---------

# 0. Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(htmlwidgets)
rm(list=ls())
```


# 1. Import all covariate data

```{r message=FALSE}
# Old code to fix everything
#bind_rows(read_csv("./data/output/OSM_SBFI2020_site_covariates.csv"),
#          read_csv("./data/output/OSM_SBFI2020_site_covariates10000.csv")) %>%
#  
#  write_csv(., file = "./data/output/OSM_SBFI2020_site_covariates.csv")


hfi <- bind_rows(read_csv("./data/raw/OSM_HFI2021_site_covariates.csv"),
                 read_csv("./data/raw/OSM_HFI2022_site_covariates.csv"))

sbfi <- bind_rows(read_csv("./data/raw/OSM_SBFI2020_site_covariates.csv")) %>%
  
  select(-lat, -long)

setdiff(sbfi$site, hfi$site)

# Read in the site/year combos. 
osm_site_year <- read_csv("./data/processed/OSM_site_coordinates_2021-2024.csv") %>%
  
  mutate(array_year = year(start_date)) %>%
  
  select(site, array_year)
```

Some 'control' sites didn't extract in the HFI because they have exactly 0 disturbance. Crazy! Let's manually add them here using the SBFI as a reference.

```{r}
hfi <- hfi %>%
  
  complete(site = sbfi$site, buffer_dist) %>%
  # Remove array
  select(-array) %>%
  # Add the correct array from the SBFI data (fixes NAs for the new sites)
  left_join(distinct(sbfi[1:2]), by = "site") %>%
  # Put array as first column
  relocate(array) %>%

  # Fill in the NA covariates that should be 0s. 
  mutate(across(4:last_col(), ~replace_na(., 0))) %>%
  
  # Manual fix since this column was weirdly a duplicate.
  # Combine the two columns and remove the one with hyphens.
  # Without this step, the next step throws an error. 
  mutate(WELL_CLEARED_NOT_DRILLED = rowSums(across(any_of(c(
    'WELL_CLEARED_NOT_DRILLED',
    'WELL-CLEARED-NOT-DRILLED'
  ))))
  ) %>%
  
  select(-contains('WELL-CLEARED-NOT-DRILLED'))

```

# 2. Tidy up the HFI data

## 2.1. Initial cleaning, add the sample year column

This is a bit janky because of the repeat samples... will want to find a better solution next year. 

```{r}
hfi <- hfi %>%
  
  set_names(tolower(names(.)) %>%
              str_replace_all(., pattern = c('-'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_') %>%  # what you want the pattern to be replaced with
              str_replace_all(., pattern = c('\\.'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_')
            ) %>%

  # What year were the arrays sampled? Used to determine recent fire and burns.
  # Currently, left joined from the deployment data frame
  left_join(osm_site_year, by = "site") #%>%
  
  # Old code from merging the ancestral arrays
  #mutate(array_year = case_when(
  #    array %in% c("WF1", "WF2", "RC") ~ 2018,
  #    array %in% c("CL") ~ 2011)
  #  )
```

```{r}
names(hfi)
```

## 1.2. Group the HFI data

We will follow the convention used by Marissa Dyck, ABMI, etc. to group variables. This code is nice and robust to whether a column is present in the data or not. 

```{r format covs}
hfi_grouped <- hfi %>% 
  
  # rename 'vegetated_edge_roads' so that we can use 'road' as keyword to group roads without including this feature
  rename(vegetated_edge_rds = vegetated_edge_roads) %>%
  
  mutate(
    # borrowpits
    borrowpits = rowSums(across(any_of(c(
      grep("borrowpit", names(.), value = TRUE),
      'dugout', 
      'lagoon', 
      'sump'
    )))),
    
    # clearings
    clearings = rowSums(across(any_of(c(
      grep("clearing", names(.), value = TRUE),
      'runway'
    )))),
    
    # cultivations
    cultivation = rowSums(across(any_of(c(
      'crop', 
      'cultivation_abandoned', 
      'fruit_vegetables',
      'rough_pasture', 
      'tame_pasture'
    )))),
    
    # industrial facilities
    facilities = rowSums(across(any_of(c(
      grep("facility", names(.), value = TRUE),
      grep("plant", names(.), value = TRUE),
      'camp_industrial', 
      'urban_industrial',
      'mill', 
      'ris_camp_industrial', 
      'ris_tank_farm', 
      'ris_utilities'
    )))),
    
    # harvest areas
    harvest_total = rowSums(across(contains('harvest_area'))),
    
    # mine areas
    mines = rowSums(across(any_of(c(
      grep("mine", names(.), value = TRUE),
      grep("tailing", names(.), value = TRUE),
      'grvl_sand_pit', 
      'peat', 
      'ris_drainage', 
      'ris_oilsands_rms',
      'ris_overburden_dump', 
      'ris_reclaim_ready', 
      'ris_soil_salvaged', 
      'ris_waste'
    )))),
    
    # railways
    railways = rowSums(across(any_of(grep("rlwy", names(.), value = TRUE)))),
    
    # reclaimed areas
    reclaimed = rowSums(across(any_of(c(
      grep("reclaimed", names(.), value = TRUE),
      'ris_soil_replaced', 
      'ris_windrow'
    )))),
    
    # recreation areas
    recreation = rowSums(across(any_of(c(
      'campground', 
      'golfcourse', 
      'greenspace', 
      'recreation'
    )))),
    
    # residential areas
    residential = rowSums(across(any_of(c(
      'country_residence', 
      'rural_residence', 
      'urban_residence'
    )))),
    
    # roads
    roads = rowSums(across(any_of(c(
      grep("road", names(.), value = TRUE),
      'transfer_station', 
      'interchange_ramp', 
      'airp_runway', 
      'ris_airp_runway'
    )))),
    
    # seismic lines
    seismic_lines_conventional = rowSums(across(any_of(c(
      'conventional_seismic',
      'pre_low_impact_seismic'
    )))),
    
    # 3D seismic lines
    seismic_lines_3D = low_impact_seismic,
    
    # transmission lines
    transmission_lines = rowSums(across(any_of(c(
      grep("transmission", names(.), value = TRUE)
    )))),
    
    # trails
    trails = rowSums(across(any_of(c(
      grep("trail", names(.), value = TRUE)
    )))),
    
    # vegetated edges
    veg_edges = rowSums(across(any_of(c(
      grep("vegetated", names(.), value = TRUE),
      'surrounding_veg'
    )))),
    
    # man-made water features
    water = rowSums(across(any_of(c(
      'canal', 
      'reservoir'
    )))),
    
    # well sites
    wells_total = rowSums(across(any_of(c(
      grep("well", names(.), value = TRUE)
    )))),
    
    # inactive well sites
    wells_inactive = rowSums(across(any_of(c(
      "well_aband"
    )))),
    
    wells_active = rowSums(across(any_of(c(
      'well_bitumen', 
      'well_bit',
      'well_gas', 
      'well_oil', 
      'well_cased',
      "well_cleared_not_confirmed", 
      "well_cleared_not_drilled",
      "well_cleared_drilled",
      "well_drilled_other",
      "ris_well",
      "well_other", 
      "well_unknown"
    )))),
    
    # remove columns that were used to create new columns to tidy the data frame
    .keep = 'unused'
  ) %>%
  
  # reorder alphabetically except array, site and buff_dist
  select(order(colnames(.))) %>% 
  
  # move the columns that aren't HFI features or landcover to the front
  relocate(array, site, buffer_dist, array_year) %>% 
  
  # reorder variables so the veg data is after all the HFI data
  relocate(starts_with('lc_class'), .after = wells_active)
```

## 1.3. Group the harvest data by year

```{r}
# Now compute row-wise harvest_0_15 and harvest_gt_15
# Loop through the sample years. This is the most efficient code I could find. 
# We want to sum the white zone AND green zone harvest for this. 
# Instead of renaming the columns, this should do the trick.
hfi_harvest <- purrr::map_dfr(unique(hfi$array_year), ~{

  # Columns to sum (based on fire year and sample year)
  cols_0_15 <- c(paste0("harvest_area_", seq(.x - 14, .x)),
                 paste0("harvest_area_white_zone_", seq(.x - 14, .x)))
                 
  cols_gt_15 <- c(paste0("harvest_area_", seq(1900, .x - 15)),
                  paste0("harvest_area_white_zone_", seq(1900, .x - 15)))
  
  # Establishment stage
  cols_10_25 <- c(paste0("harvest_area_", seq(.x - 24, .x - 10)),
                  paste0("harvest_area_white_zone_", seq(.x - 24, .x - 10)))
  
  # Aggrandation phase
  cols_gt_25 <- c(paste0("harvest_area_", seq(1900, .x - 25)),
                  paste0("harvest_area_white_zone_", seq(1900, .x - 25)))
  
  # By decade
  cols_0_10 <- c(paste0("harvest_area_", seq(.x - 9, .x)),
                 paste0("harvest_area_white_zone_", seq(.x - 9, .x)))
  
  cols_10_20 <- c(paste0("harvest_area_", seq(.x - 19, .x - 10)),
                  paste0("harvest_area_white_zone_", seq(.x - 19, .x - 10)))
  
  cols_20_30 <- c(paste0("harvest_area_", seq(.x - 29, .x - 20)),
                  paste0("harvest_area_white_zone_", seq(.x - 29, .x - 20)))
  
  cols_30_40 <- c(paste0("harvest_area_", seq(.x - 39, .x - 30)),
                  paste0("harvest_area_white_zone_", seq(.x - 39, .x - 30)))
  
  cols_gt_40 <- c(paste0("harvest_area_", seq(1900, .x - 40)),
                  paste0("harvest_area_white_zone_", seq(1900, .x - 40)))

  
  hfi %>%
    filter(array_year == .x) %>%
    mutate(
      
      # Sum them
      harvest_0_15 = rowSums(across(any_of(
        cols_0_15
      ))),
      
      harvest_gt_15 = rowSums(across(any_of(
        cols_gt_15
      ))),
      
      harvest_gt_25 = rowSums(across(any_of(
        cols_gt_25
      ))),
      
      harvest_0_10 = rowSums(across(any_of(
        cols_0_10
      ))),
      
      harvest_10_25 = rowSums(across(any_of(
        cols_10_25
      ))),
      
      harvest_10_20 = rowSums(across(any_of(
        cols_10_20
      ))),
      harvest_20_30 = rowSums(across(any_of(
        cols_20_30
      ))),
      harvest_30_40 = rowSums(across(any_of(
        cols_30_40
      ))),
      harvest_gt_40 = rowSums(across(any_of(
        cols_gt_40
      )))
      
    ) %>%
    
  # move the columns that aren't HFI features or landcover to the front
  select(array, site, buffer_dist, array_year, 
         harvest_0_15, 
         harvest_gt_15,
         
         harvest_0_10,
         harvest_10_25,
         harvest_gt_25,
         
         harvest_10_20,
         harvest_20_30,
         harvest_30_40,
         harvest_gt_40)
    
  }
  )
```

## 1.4. Add the harvest and do some additional grouping

```{r}
hfi_final <- hfi_grouped %>% 
  
  left_join(hfi_harvest, by = c("array", "site", "buffer_dist", "array_year")) %>%

  mutate(
    # create column osm_industrial following Marissa's convention
    osm_industrial = 
      borrowpits +
      clearings +
      facilities +
      mines,
    
    # Lump popeline and transmission line
    pipe_trans =
      pipeline +
      transmission_lines,
    
    # Lump seismic
    seismic_total = 
      seismic_lines_conventional + 
      seismic_lines_3D,
    
    # Cumulative site disturbance without harvest
    cfi =
      osm_industrial +
      wells_total + 
      seismic_total + 
      pipe_trans +
      trails +
      roads + 
      railways,
    
    # Cumulative site disturbance with veg edges
    cfi_with_vegedges = 
      cfi + 
      veg_edges,
    
        # Cumulative site disturbance with harvest
    cfi_with_harvest = 
      cfi_with_vegedges + 
      harvest_0_15,

    ) %>%
  
  # remove other features we don't need
  select(!any_of(c(
    'cfo',
    'cultivation',
    'reclaimed',
    'recreation',
    'residential',
    'water',
    'borrowpits',
    'clearings',
    'facilities',
    'mines',
    'pipeline',
    'transmission_lines',
    'landfills',
    'other_linear_features',
    'woody_vegetation_removal'
  ))) %>%
  
  relocate(array, site, buffer_dist, array_year, sort(names(.)))

names(hfi_final)
```

Looks good!!

# 3. Tidy up the forest inventory data

## 3.1. Clean up and add sample year column

```{r}
sbfi <- sbfi %>%
  
  set_names(tolower(names(.)) %>%
              str_replace_all(., pattern = c('-'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_') %>%  # what you want the pattern to be replaced with
              str_replace_all(., pattern = c('\\.'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_')
            ) %>%

  # What year were the arrays sampled? Used to determine recent fire and burns.
  # Currently, left joined from the deployment data frame
  left_join(osm_site_year, by = "site") %>%
  
  # Old code from merging the ancestral arrays
  #mutate(array_year = case_when(
  #    array %in% c("WF1", "WF2", "RC") ~ 2018,
  #    array %in% c("CL") ~ 2011)
  #  )
  
  mutate(
    # fire
    fire_total = rowSums(across(contains('fire_pct')))
    )

  
names(sbfi)
```

## 3.2. Calculate fire by year columns

```{r}
# Now compute row-wise fire_0_15 and fire_gt_15
# Loop through the sample years. This is the most efficient code I could find. 
sbfi_fire <- purrr::map_dfr(unique(sbfi$array_year), ~{
  
  # Columns to sum (based on fire year and sample year)
  cols_0_15 <- paste0("fire_pct_", seq(.x - 14, .x))
  cols_gt_15 <- paste0("fire_pct_", seq(1900, .x - 15))
  
  # By decade
  cols_0_10 <- paste0("fire_pct_", seq(.x - 9, .x))
  cols_10_20 <- paste0("fire_pct_", seq(.x - 19, .x - 10))
  cols_20_30 <- paste0("fire_pct_", seq(.x - 29, .x - 20))
  cols_30_40 <- paste0("fire_pct_", seq(.x - 39, .x - 30))
  cols_gt_40 <- paste0("fire_pct_", seq(1900, .x - 40))
  
  # Establishment and aggrandization stages:
  cols_10_25 <- paste0("fire_pct_", seq(.x - 24, .x - 10))
  cols_gt_25 <- paste0("fire_pct_", seq(1900, .x - 25))


  sbfi %>%
    filter(array_year == .x) %>%
    mutate(
      
      # Sum them
      fire_0_15 = rowSums(across(any_of(
        cols_0_15
      ))),
      
      fire_gt_15 = rowSums(across(any_of(
        cols_gt_15
      ))),
      
      fire_gt_25 = rowSums(across(any_of(
        cols_gt_25
      ))),
      
      fire_10_25 = rowSums(across(any_of(
        cols_10_25
      ))),
      
      fire_0_10 = rowSums(across(any_of(
        cols_0_10
      ))),
      fire_10_20 = rowSums(across(any_of(
        cols_10_20
      ))),
      fire_20_30 = rowSums(across(any_of(
        cols_20_30
      ))),
      fire_30_40 = rowSums(across(any_of(
        cols_30_40
      ))),
      fire_gt_40 = rowSums(across(any_of(
        cols_gt_40
      )))
      
    ) %>%
    
  # move the columns that aren't HFI features or landcover to the front
  select(array, site, buffer_dist, array_year, 
         fire_0_15, 
         fire_gt_15,
         fire_0_10,
         fire_10_25,
         fire_gt_25,
         fire_10_20,
         fire_20_30,
         fire_30_40,
         fire_gt_40)
  }
  )
```

## 3.3. Finalize and give the species a more meaningful name

```{r}
sbfi_final <- sbfi %>% 
  
  # Some columns we don't want
  select(-contains("fire"),
         -contains("age"), 
         -lc_rock_rubble,
         -lc_snow_ice,
         -lc_bryoids,
         -lc_exposed_barren,
         # We'll get harvest from the ABMI data instead
         -contains("harvest"),
         # We DO want fire total though. Overrides previous removal.
         fire_total) %>%
  
  # Join the fire data
  left_join(sbfi_fire, by = c("array", "site", "buffer_dist", "array_year")) %>%

  relocate(array, site, buffer_dist, array_year, sort(names(.))) %>%
  
  # Nicer species column names
  rename_with(
    ~ paste0("pct_", str_remove(., "_pct_of_treed")),
    .cols = contains("_pct_of_treed")
  )

names(sbfi_final)
```



# 5. Put it all together:

## 5.1. Import lat/long and UTM 12N coordinates

```{r}
# Clean up this similar, old file
rm(osm_site_year)

# read in the 'site coordinate' files. Purrr works for now. 
coords <- read_csv("./data/processed/OSM_site_coordinates_2021-2024.csv")
``` 

## 5.2. Append everything and final inspection

```{r}
# Tidy up and final check
covs <- hfi_final %>%
  
  left_join(sbfi_final, by = c("array", "site", "buffer_dist", "array_year")) %>%

  left_join(coords, by = c('array', 'site')) %>%
  
  relocate(array, site, array_year, start_date, end_date, lat, long, easting_12n, northing_12n, buffer_dist) %>%
  
  # Remove some columns. Should have been done in an earlier step but just noticed it now...
  select(-camera_failure_details, -deployment_id)

summary(covs)
head(covs)
```

Also append the NDVI:

```{r}
ndvi <- read_csv("./data/raw/OSM_MODISNDVI_site_covariates.csv")

covs <- covs %>%
  
  left_join(ndvi, by = c("site", "array", "buffer_dist", "lat", "long")) %>%
  
  # Remove some columns. Should have been done in an earlier step but just noticed it now...
  select(-camera_failure_details, -deployment_id)
```


## 5.4. Save the data:

```{r}
write_csv(covs, "./data/processed/OSM_site_covariates_2021-2024.csv")
```



End script




