---
title: "Extract landcover covariates"
author: "Aidan Brushett"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: default
    toc: yes
    toc_float: yes
---

```{r setup, include=TRUE}
# This is set to false just for quick knitting
knitr::opts_chunk$set(echo = T,
                      eval = T)
```

```{r}
library(sf) # for points and buffering
library(tidyverse)
rm(list = ls())
```

PLEASE READ ME!!!

Before using R to extract the covariates, I selected the portion of the datasets around the sites using the "Select Features by Area" tool. I DO NOT RECOMMMEND CLIPPING the files as this caused loss of features. I did this in QGIS because R was very slow for processing the full layers, although possible if you have a more powerful computer. 

The steps for this are (Follow steps 1 to 3 if subsetting the layer is required): 
1. Use the "Select Features by Area" tool to select a portion of the layer and export as a new shapefile (see note about CRS below). Make 7 km buffers around the site to ensure you select a large enough portion of the layer (change this if your largest buffers will be larger than 5km.
2. Manually look at the layers to make sure all features were selected (important for HFI). Zoom in and compare the subsetted and the original layers.
3. Ensure that the camera sites align in the correct locations (I used lakes/towns to double-check and visually reference this, just to be sure CRS are all good).
4. Reproject the layer to the appropriate UTM Zone. 
5. Follow the below script in R with the subsetted/reprojected data.
6. QAQC in QGIS or ArcGIS: compare the output dataframe and the original layer. I created a 250 m and 1000 m buffers and looked a multiple sites in detail to ensure each feature type was represented in the output .csv for those buffers. 


!!! IMPORTANT NOTE !!! Check the validity of all layer CRS BEFORE extracting covariates in R. READ BELOW. 


The ABMI layers are in NAD83 / Alberta 10-TM (Forest) when downloaded but extracting in UTM Zones is slightly more accurate due to the smaller area and the position of the central meridian. If sites are in western Alberta, UTM Zone 11 will need to be used (make sure to check a map of UTM Zones and choose the most suitable one). Once you select the area of interest, save the shape file as a new shape file with the UTM Zone CRS. 

# 1. Read in the raw site coordinates:

I'll read in all the sites into a single file first: 

```{r}
sites_osm <- read_csv("./data/input/OSM_site_coordinates_2021-2024.csv") %>%
  
  st_as_sf(., coords = c("easting_12n", "northing_12n"), crs = 26912)
```

We will also subset based on survey year, since 2021 and 2022 use separate human footprint databases from ABMI. 

```{r}
# Sampled 2023 onwards
sites_osm_2023 <- sites_osm %>% 
  
  filter(str_detect(array_visit, "2023|2024|2025"))

# Sampled 2022
sites_osm_2022 <- sites_osm %>% 
  
  filter(str_detect(array_visit, "2022"))

# Sampled 2021 
sites_osm_2021 <- sites_osm %>% 
  
  filter(str_detect(array_visit, "2021"))
```


# 2. Forest inventory data

## 2.1. Load, clip the forest inventory data

We can do this for all sites at once (i.e. not split by survey year, since it's the same dataset for all arrays). First we'll buffer the camera sites so we can tell what 

```{r}
#For combining multiple sf objects into 1 (not needed currently other than for renaming to `sites`)
sites <- bind_rows(sites_osm) #, sites_christina)

# Check it out on a map
plot(sites)

# buffer the sites so that we can filter the landcover dataset and reduce file size
# An 11km buffer will encompass all features for the buffer radii that we are extracting. 
sites_buffered <- st_buffer(sites, dist = 11000) %>%
  st_union()

# check it out on a map
plot(sites_buffered[1])
```


# 3. Human footprint data 2021 -- for OSM sites in 2021 ONLY

Future years are extracted using the 2022 HFI for now. 

## 3.1. Import data

```{r}
# Camera locations (point)
sites <- sites_osm_2021
# buffer sites so that we can filter the landcover data and reduce file size
sites_buffered <- st_buffer(sites, dist = 11000) %>%
  
  st_union()

plot(sites_buffered[1])

# Human footprint index:
# load human features layer (will be slow)
#st_layers("./maps/HFI2021_clipped.gdb")

hfi <- st_read("C:/databases/HFI2021_boreal_plains/HFI_2021_clipped.shp") %>%
  
  st_transform(., crs = 26912)


#make sure it's all projected correctly 
st_crs(hfi, describe = TRUE)
st_crs(sites, describe = TRUE)
#plot(hfi[1])

# clean up
rm(sites_buffered)
gc()

# double check that they are all format "sf" and "dataframe"
str(sites)
str(hfi)
```

## 3.2. Extract variables

```{r message=FALSE, warning=FALSE}
# Extract HFI (Human Footprint Inventory) covariates using site buffers
hfi_metrics <- hfi %>%
  select(OBJECTID, FEATURE_TY, YEAR) # Keep only relevant columns

sites_hfi_metrics <- list() # List to store output

start_time <- Sys.time() # Start timing the process

# Loop over each site in the 'sites' object
for(i in 1:nrow(sites)) {
  
  progress <- (i - 1) / nrow(sites) # Track progress
  
  site <- sites[i, ] # Extract the current site
  
  # Print progress to console with estimated time remaining
  cat("\r Extracting site ", i, " of ", nrow(sites), " (", site$site, "): ", 
      round(as.numeric(difftime(Sys.time(), start_time, units = "hours")) * (1 - progress) / progress, 2), 
      " hours remaining        ", sep = "")
  
  # Extract metrics for the current site across buffer distances
  sites_hfi_metrics[[i]] <-
    
    purrr::map_dfr(c(50, 100, 200, 250, 300, 400, seq(500, 5000, by = 250), 10000), ~{
      
      site_buffer <- st_buffer(site, .x) # Create buffer around site
      
      st_intersection(hfi_metrics, site_buffer) %>% # Intersect with HFI layer
        
        # Calculate areas and percent cover
        mutate(feature_area = st_area(.) %>%
                 as.numeric(.), # Area of each intersected HFI polygon
               
               total_area = st_area(site_buffer) %>%
                 as.numeric(.), # Total area of buffer (unlike landcover, we use full area)
               
               percent_area = feature_area / total_area) %>% # Proportion of buffer occupied
        
        st_drop_geometry(.) %>% # Drop geometry for efficiency
        
        # Differentiate harvested areas by year
        mutate(FEATURE_TY = ifelse(str_detect(FEATURE_TY, "HARVEST-AREA"),
                                   paste0(FEATURE_TY, "-", YEAR),
                                   FEATURE_TY)) %>%
        
        # Group by array/site/feature type to summarize percent cover
        group_by(array, site, array_visit, FEATURE_TY) %>%
        
        summarize(PCT_COVER = sum(percent_area),
                  buffer_dist = .x) %>%
        
        # Pivot to wide format: one row per site-buffer combo, features as columns
        pivot_wider(names_from = FEATURE_TY,
                    values_from = PCT_COVER,
                    values_fill = 0)
    })
  
  # Clean up memory after each iteration
  rm(site)
  gc()
}

save(sites_hfi_metrics, file = "./data/output/OSM_HFI2021_site_covariates.RData")
```


## 3.3. Save results

```{r message=FALSE}
# Some extra tidying. We need to deal with missing buffers/sites when there were no human features. 
sites_hfi_metrics_df <- bind_rows(sites_hfi_metrics) %>%
  
  ungroup() %>%

  relocate(array, site, array_visit, buffer_dist, sort(names(.))) #%>%
  
  #select(-array) %>%
  
  #complete(site, buffer_dist) %>%
  
  # Create some extra columns
  #separate_wider_delim(site,
  #                     delim = '_',
  #                     names = c('array', 'camera'),
  #                     cols_remove = FALSE) %>%
  
  #select(-camera) %>%
  
  # Replace NAs introduced by "complete"
  #mutate(across(4:last_col(), ~replace(., is.na(.), 0)))

write_csv(sites_hfi_metrics_df, "./data/output/OSM_HFI2021_site_covariates.csv")
```

***





# 4. Human footprint data 2022 -- for OSM sites in 2022 ONWARDS

Previous years are extracted using the 2021 HFI for now. 

## 4.1. Import data

```{r}
# Camera locations (point)
sites <- sites_osm_2022
# buffer sites so that we can filter the landcover data and reduce file size
sites_buffered <- st_buffer(sites, dist = 11000) %>%
  
  st_union()

plot(sites_buffered[1])

# Human footprint index:
# load human features layer (will be slow)
#st_layers("./maps/HFI2021_clipped.gdb")

hfi <- st_read("C:/databases/HFI2022_boreal_plains/HFI_2022_clipped.shp") %>%
  
  st_transform(., crs = 26912)


#make sure it's all projected correctly 
st_crs(hfi, describe = TRUE)
st_crs(sites, describe = TRUE)
#plot(hfi[1])

# clean up
rm(sites_buffered)
gc()

# double check that they are all format "sf" and "dataframe"
str(sites)
str(hfi)
```
Align the timber harvest names so that they correspond to the harvest names in the 2021 HFI. I don't know why they changed this...

```{r}
sort(unique(hfi$FEATURE_TY))

hfi <- hfi %>%
  
  mutate(FEATURE_TY = case_when(
    FEATURE_TY == "TIMBER-HARVEST-GREEN-AREA" ~ "HARVEST-AREA",
    FEATURE_TY == "TIMBER-HARVEST-WHITE-AREA" ~ "HARVEST-AREA-WHITE-ZONE",
    TRUE ~ FEATURE_TY))

sort(unique(hfi$FEATURE_TY))

```


## 4.2. Extract variables

```{r message=FALSE, warning=FALSE}
# Extract HFI (Human Footprint Inventory) covariates using site buffers
hfi_metrics <- hfi %>%
  select(OBJECTID, FEATURE_TY, YEAR) # Keep only relevant columns

sites_hfi_metrics <- list() # List to store output

start_time <- Sys.time() # Start timing the process

# Loop over each site in the 'sites' object
for(i in 1:nrow(sites)) {
  
  progress <- (i - 1) / nrow(sites) # Track progress
  
  site <- sites[i, ] # Extract the current site
  
  # Print progress to console with estimated time remaining
  cat("\r Extracting site ", i, " of ", nrow(sites), " (", site$site, "): ", 
      round(as.numeric(difftime(Sys.time(), start_time, units = "hours")) * (1 - progress) / progress, 2), 
      " hours remaining        ", sep = "")
  
  # Extract metrics for the current site across buffer distances
  sites_hfi_metrics[[i]] <-
    
    purrr::map_dfr(c(50, 100, 200, 250, 300, 400, seq(500, 5000, by = 250), 10000), ~{
      
      site_buffer <- st_buffer(site, .x) # Create buffer around site
      
      st_intersection(hfi_metrics, site_buffer) %>% # Intersect with HFI layer
        
        # Calculate areas and percent cover
        mutate(feature_area = st_area(.) %>%
                 as.numeric(.), # Area of each intersected HFI polygon
               
               total_area = st_area(site_buffer) %>%
                 as.numeric(.), # Total area of buffer (unlike landcover, we use full area)
               
               percent_area = feature_area / total_area) %>% # Proportion of buffer occupied
        
        st_drop_geometry(.) %>% # Drop geometry for efficiency
        
        # Differentiate harvested areas by year
        mutate(FEATURE_TY = ifelse(str_detect(FEATURE_TY, "HARVEST-AREA"),
                                   paste0(FEATURE_TY, "-", YEAR),
                                   FEATURE_TY)) %>%
        
        # Group by array/site/feature type to summarize percent cover
        group_by(array, site, array_visit, FEATURE_TY) %>%
        
        summarize(PCT_COVER = sum(percent_area),
                  buffer_dist = .x) %>%
        
        # Pivot to wide format: one row per site-buffer combo, features as columns
        pivot_wider(names_from = FEATURE_TY,
                    values_from = PCT_COVER,
                    values_fill = 0)
    })
  
  # Clean up memory after each iteration
  rm(site)
  gc()
}

save(sites_hfi_metrics, file = "./data/output/OSM_HFI2022_site_covariates.RData")
```


## 4.3. Save results

```{r message=FALSE}
# Some extra tidying. We need to deal with missing buffers/sites when there were no human features. 
sites_hfi_metrics_df <- bind_rows(sites_hfi_metrics) %>%
  
  ungroup() %>%

  relocate(array, site, array_visit, buffer_dist, sort(names(.))) #%>%
  
  #select(-array) %>%
  
  #complete(site, buffer_dist) %>%
  
  # Create some extra columns
  #separate_wider_delim(site,
  #                     delim = '_',
  #                     names = c('array', 'camera'),
  #                     cols_remove = FALSE) %>%
#  
  #select(-camera) %>%
  
  # Replace NAs introduced by "complete"
  #mutate(across(4:last_col(), ~replace(., is.na(.), 0)))

write_csv(sites_hfi_metrics_df, "./data/output/OSM_HFI2022_site_covariates.csv")
```

***







# 5. Human footprint data 2023 -- for OSM sites in 2023 ONWARDS

Previous years are extracted using the 2021 and 2022 HFI for now. Age attribution gets better every year in the dataset, and we should be able to just use a single HFI layer next year (and filter by `YEAR` column)

## 5.1. Import data

```{r}
# Camera locations (point)
sites <- sites_osm_2023
# buffer sites so that we can filter the landcover data and reduce file size
sites_buffered <- st_buffer(sites, dist = 11000) %>%
  
  st_union()

plot(sites_buffered[1])

# Human footprint index:
# load human features layer (will be slow)
#st_layers("./maps/HFI2021_clipped.gdb")

hfi <- st_read("C:/databases/HFI2023_boreal_plains/HFI_2023_clipped.shp") %>%
  
  st_transform(., crs = 26912)


#make sure it's all projected correctly 
st_crs(hfi, describe = TRUE)
st_crs(sites, describe = TRUE)
#plot(hfi[1])

# clean up
rm(sites_buffered)
gc()

# double check that they are all format "sf" and "dataframe"
str(sites)
str(hfi)
```
Align the timber harvest names so that they correspond to the harvest names in the 2021 HFI. I don't know why they changed this...

```{r}
sort(unique(hfi$FEATURE_TY))

hfi <- hfi %>%
  
  mutate(FEATURE_TY = case_when(
    FEATURE_TY == "TIMBER-HARVEST-GREEN-AREA" ~ "HARVEST-AREA",
    FEATURE_TY == "TIMBER-HARVEST-WHITE-AREA" ~ "HARVEST-AREA-WHITE-ZONE",
    FEATURE_TY == "Seismic - Legacy" ~ "CONVENTIONAL-SEISMIC",
    FEATURE_TY == "Seismic - LIS" ~ "LOW-IMPACT-SEISMIC",
    FEATURE_TY == "Other - Linear" ~ "OTHER-LINEAR-FEATURES",
    TRUE ~ FEATURE_TY))

sort(unique(hfi$FEATURE_TY))

```


## 5.2. Extract variables

```{r message=FALSE, warning=FALSE}
# Extract HFI (Human Footprint Inventory) covariates using site buffers
hfi_metrics <- hfi %>%
  select(OBJECTID, FEATURE_TY, YEAR) # Keep only relevant columns

sites_hfi_metrics <- list() # List to store output

start_time <- Sys.time() # Start timing the process

# Loop over each site in the 'sites' object
for(i in 1:nrow(sites)) {
  
  progress <- (i - 1) / nrow(sites) # Track progress
  
  site <- sites[i, ] # Extract the current site
  
  # Print progress to console with estimated time remaining
  cat("\r Extracting site ", i, " of ", nrow(sites), " (", site$site, "): ", 
      round(as.numeric(difftime(Sys.time(), start_time, units = "hours")) * (1 - progress) / progress, 2), 
      " hours remaining        ", sep = "")
  
  # Extract metrics for the current site across buffer distances
  sites_hfi_metrics[[i]] <-
    
    purrr::map_dfr(c(50, 100, 200, 250, 300, 400, seq(500, 5000, by = 250), 10000), ~{
      
      site_buffer <- st_buffer(site, .x) # Create buffer around site
      
      st_intersection(hfi_metrics, site_buffer) %>% # Intersect with HFI layer
        
        # Calculate areas and percent cover
        mutate(feature_area = st_area(.) %>%
                 as.numeric(.), # Area of each intersected HFI polygon
               
               total_area = st_area(site_buffer) %>%
                 as.numeric(.), # Total area of buffer (unlike landcover, we use full area)
               
               percent_area = feature_area / total_area) %>% # Proportion of buffer occupied
        
        st_drop_geometry(.) %>% # Drop geometry for efficiency
        
        # Differentiate harvested areas by year
        mutate(FEATURE_TY = ifelse(str_detect(FEATURE_TY, "HARVEST-AREA"),
                                   paste0(FEATURE_TY, "-", YEAR),
                                   FEATURE_TY)) %>%
        
        # Group by array/site/feature type to summarize percent cover
        group_by(array, site, array_visit, FEATURE_TY) %>%
        
        summarize(PCT_COVER = sum(percent_area),
                  buffer_dist = .x) %>%
        
        # Pivot to wide format: one row per site-buffer combo, features as columns
        pivot_wider(names_from = FEATURE_TY,
                    values_from = PCT_COVER,
                    values_fill = 0)
    })
  
  # Clean up memory after each iteration
  rm(site)
  gc()
}

save(sites_hfi_metrics, file = "./data/output/OSM_HFI2023_site_covariates.RData")
```


## 5.3. Save results

```{r message=FALSE}
# Some extra tidying. We need to deal with missing buffers/sites when there were no human features. 
sites_hfi_metrics_df <- bind_rows(sites_hfi_metrics) %>%
  
  ungroup() %>%

  relocate(array, site, array_visit, buffer_dist, sort(names(.))) #%>%
  
#  select(-array) %>%
#  
#  complete(site, buffer_dist) %>%
#  
#  # Create some extra columns
#  separate_wider_delim(site,
#                       delim = '_',
#                       names = c('array', 'camera'),
#                       too_many = "merge",
#                       cols_remove = FALSE) %>%
#  
#  select(-camera) %>%
#  
#  # Replace NAs introduced by "complete"
#  mutate(across(4:last_col(), ~replace(., is.na(.), 0)))

write_csv(sites_hfi_metrics_df, "./data/output/OSM_HFI2023_site_covariates.csv")
```

***

