---
title: "Format covariates"
author: "Aidan Brushett"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: default
    toc: yes
    toc_float: yes
---

# Before you begin

Aidan Brushett
Ph.D. Student
University of Victoria    
School of Environmental Studies     
Email: [aidanbrushett@uvic.ca](aidanbrushett@uvic.ca)

---------

# 0. Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(htmlwidgets)
rm(list=ls())
```


# 1. Import all covariate data

FOR 2026/27: THE ARRAY_VISIT COLUMN HAS BEEN ADDED IN THE EXTRACTION SCRIPTS TO ACCOUNT FOR REPEAT SAMPLES/MULTIPLE YEARS WITH THE SAME SITE. 
Fixes to this script must include:
- joining the data based on the `array_visit` column (I did it manually for 2025/26 but this is the last time that will work). 
- accounting for missing values/rows in the HFI. This is hard to do now in Script 1x. because of the `array_visit` column. This could all be accomplished by merging to the SBFI, which should include all sites and all buffers (no missing valueS) 
  - I think the two chunks below already handle this, just need to confirm

```{r message=FALSE}
hfi <- bind_rows(read_csv("./data/output/OSM_HFI2021_site_covariates.csv"),
                 read_csv("./data/output/OSM_HFI2022_site_covariates.csv"),
                 read_csv("./data/output/OSM_HFI2023_site_covariates.csv"))

sbfi <- read_csv("./data/output/OSM_SBFI2020_site_covariates.csv")

fire <- read_csv("./data/output/OSM_ABWILDFIRE_site_covariates.csv")

daymet <- read_csv("./data/output/OSM_DAYMET_annual_site_covariates.csv")

ndvi <- read_csv("./data/output/OSM_MODISNDVI_site_covariates.csv")
```


!!!! TEMPORARY FIX (WON'T BE NEEDED THE NEXT TIME THE SCRIPT IS RUN)
```{r message=FALSE}
#!!!! TEMPORARY FIX (WON'T BE NEEDED THE NEXT TIME THE SCRIPT IS RUN)
#sites_osm <- read_csv("./data/input/OSM_site_coordinates_2021-2024.csv") %>%
#  
#  select(array, site, array_visit)#
#
#
#sbfi <- left_join(sbfi, sites_osm, by = c("array", "site")) %>%
#  
#  relocate(array, site, array_visit)

# OVERWRITE
#write_csv(sbfi, "./data/output/OSM_SBFI2020_site_covariates.csv")
```

Some 'control' sites didn't extract in the HFI because they have exactly 0 disturbance. Crazy! Let's manually add them here using the SBFI as a reference.

```{r}
hfi <- hfi %>%
  
  complete(site = sbfi$site, buffer_dist) %>%
  # Remove array and array_visit, they'll be re-joined from the SBFI dataset
  select(-any_of(c("array", "array_visit"))) %>%
  # Add the correct array and array_visit from the SBFI data (fixes NAs for the new sites)
  left_join(distinct(sbfi[1:3]), by = "site") %>%
  # Put array as first column
  relocate(array, site, array_visit) %>%

  # Fill in the NA covariates that should be 0s. 
  mutate(across(5:last_col(), ~replace_na(., 0))) %>%
  
  # Manual fix since this column was weirdly a duplicate.
  # Combine the two columns and remove the one with hyphens.
  # Without this step, the next step throws an error. 
  mutate(WELL_CLEARED_NOT_DRILLED = rowSums(across(any_of(c(
    'WELL_CLEARED_NOT_DRILLED',
    'WELL-CLEARED-NOT-DRILLED'
  ))))
  ) %>%
  
  select(-contains('WELL-CLEARED-NOT-DRILLED'))

```

# 2. Tidy up the HFI data

## 2.1. Initial cleaning, add the sample year column

```{r}
hfi <- hfi %>%
  
  set_names(tolower(names(.)) %>%
              str_replace_all(., pattern = c('-'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_') %>%  # what you want the pattern to be replaced with
              str_replace_all(., pattern = c('\\.'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_')
            ) %>%

  # What year were the arrays sampled? Used to determine recent fire and burns.
  mutate(array_year = as.numeric(str_extract(array_visit, "\\d+$"))) %>%
  relocate(array_year, .after=array_visit)
```

```{r}
names(hfi)
```

## 1.2. Group the HFI data

We will follow the convention used by Marissa Dyck, ABMI, etc. to group variables. This code is nice and robust to whether a column is present in the data or not. 

```{r format covs}
hfi_grouped <- hfi %>% 
  
  # rename 'vegetated_edge_roads' so that we can use 'road' as keyword to group roads without including this feature
  rename(vegetated_edge_rds = vegetated_edge_roads) %>%
  
  mutate(
    # borrowpits
    borrowpits = rowSums(across(any_of(c(
      grep("borrowpit", names(.), value = TRUE),
      'dugout', 
      'lagoon', 
      'sump'
    )))),
    
    # clearings
    clearings = rowSums(across(any_of(c(
      grep("clearing", names(.), value = TRUE),
      'runway'
    )))),
    
    # cultivations
    cultivation = rowSums(across(any_of(c(
      'crop', 
      'cultivation_abandoned', 
      'fruit_vegetables',
      'rough_pasture', 
      'tame_pasture'
    )))),
    
    # industrial facilities
    facilities = rowSums(across(any_of(c(
      grep("facility", names(.), value = TRUE),
      grep("plant", names(.), value = TRUE),
      'camp_industrial', 
      'urban_industrial',
      'mill', 
      'ris_camp_industrial', 
      'ris_tank_farm', 
      'ris_utilities'
    )))),
    
    # harvest areas
    harvest_total = rowSums(across(contains('harvest_area'))),
    
    # mine areas
    mines = rowSums(across(any_of(c(
      grep("mine", names(.), value = TRUE),
      grep("tailing", names(.), value = TRUE),
      'grvl_sand_pit', 
      'peat', 
      'ris_drainage', 
      'ris_oilsands_rms',
      'ris_overburden_dump', 
      'ris_reclaim_ready', 
      'ris_soil_salvaged', 
      'ris_waste'
    )))),
    
    # railways
    railways = rowSums(across(any_of(grep("rlwy", names(.), value = TRUE)))),
    
    # reclaimed areas
    reclaimed = rowSums(across(any_of(c(
      grep("reclaimed", names(.), value = TRUE),
      'ris_soil_replaced', 
      'ris_windrow'
    )))),
    
    # recreation areas
    recreation = rowSums(across(any_of(c(
      'campground', 
      'golfcourse', 
      'greenspace', 
      'recreation'
    )))),
    
    # residential areas
    residential = rowSums(across(any_of(c(
      'country_residence', 
      'rural_residence', 
      'urban_residence'
    )))),
    
    # roads
    roads = rowSums(across(any_of(c(
      grep("road", names(.), value = TRUE),
      'transfer_station', 
      'interchange_ramp', 
      'airp_runway', 
      'ris_airp_runway'
    )))),
    
    # seismic lines
    seismic_lines_conventional = rowSums(across(any_of(c(
      'conventional_seismic',
      'pre_low_impact_seismic'
    )))),
    
    # 3D seismic lines
    seismic_lines_3D = low_impact_seismic,
    
    # transmission lines
    transmission_lines = rowSums(across(any_of(c(
      grep("transmission", names(.), value = TRUE)
    )))),
    
    # trails
    trails = rowSums(across(any_of(c(
      grep("trail", names(.), value = TRUE)
    )))),
    
    # vegetated edges
    veg_edges = rowSums(across(any_of(c(
      grep("vegetated", names(.), value = TRUE),
      'surrounding_veg'
    )))),
    
    # man-made water features
    water = rowSums(across(any_of(c(
      'canal', 
      'reservoir'
    )))),
    
    # well sites
    wells_total = rowSums(across(any_of(c(
      grep("well", names(.), value = TRUE)
    )))),
    
    # inactive well sites
    wells_inactive = rowSums(across(any_of(c(
      "well_aband"
    )))),
    
    wells_active = rowSums(across(any_of(c(
      'well_bitumen', 
      'well_bit',
      'well_gas', 
      'well_oil', 
      'well_cased',
      "well_cleared_not_confirmed", 
      "well_cleared_not_drilled",
      "well_cleared_drilled",
      "well_drilled_other",
      "ris_well",
      "well_other", 
      "well_unknown"
    )))),
    
    # remove columns that were used to create new columns to tidy the data frame
    .keep = 'unused'
  ) %>%
  
  # reorder alphabetically except array, site and buff_dist
  select(order(colnames(.))) %>% 
  
  # move the columns that aren't HFI features or landcover to the front
  relocate(array, site, buffer_dist, array_year, array_visit) %>% 
  
  # reorder variables so the veg data is after all the HFI data
  relocate(starts_with('lc_class'), .after = wells_active)
```

## 1.3. Group the harvest data by year

```{r}
# Now compute row-wise harvest_0_15 and harvest_gt_15
# Loop through the sample years. This is the most efficient code I could find. 
# We want to sum the white zone AND green zone harvest for this. 
# Instead of renaming the columns, this should do the trick.
hfi_harvest <- purrr::map_dfr(unique(hfi$array_year), ~{

  # Columns to sum (based on fire year and sample year)
  cols_0_15 <- c(paste0("harvest_area_", seq(.x - 14, .x)),
                 paste0("harvest_area_white_zone_", seq(.x - 14, .x)))
                 
  cols_gt_15 <- c(paste0("harvest_area_", seq(1900, .x - 15)),
                  paste0("harvest_area_white_zone_", seq(1900, .x - 15)))
  
  # Establishment stage
  cols_10_25 <- c(paste0("harvest_area_", seq(.x - 24, .x - 10)),
                  paste0("harvest_area_white_zone_", seq(.x - 24, .x - 10)))
  
  # Aggrandation phase
  cols_gt_25 <- c(paste0("harvest_area_", seq(1900, .x - 25)),
                  paste0("harvest_area_white_zone_", seq(1900, .x - 25)))
  
  # By decade
  cols_0_10 <- c(paste0("harvest_area_", seq(.x - 9, .x)),
                 paste0("harvest_area_white_zone_", seq(.x - 9, .x)))
  
  cols_10_20 <- c(paste0("harvest_area_", seq(.x - 19, .x - 10)),
                  paste0("harvest_area_white_zone_", seq(.x - 19, .x - 10)))
  
  cols_20_30 <- c(paste0("harvest_area_", seq(.x - 29, .x - 20)),
                  paste0("harvest_area_white_zone_", seq(.x - 29, .x - 20)))
  
  cols_30_40 <- c(paste0("harvest_area_", seq(.x - 39, .x - 30)),
                  paste0("harvest_area_white_zone_", seq(.x - 39, .x - 30)))
  
  cols_gt_40 <- c(paste0("harvest_area_", seq(1900, .x - 40)),
                  paste0("harvest_area_white_zone_", seq(1900, .x - 40)))

  
  hfi %>%
    filter(array_year == .x) %>%
    mutate(
      
      # Sum them
      harvest_0_15 = rowSums(across(any_of(
        cols_0_15
      ))),
      
      harvest_gt_15 = rowSums(across(any_of(
        cols_gt_15
      ))),
      
      harvest_gt_25 = rowSums(across(any_of(
        cols_gt_25
      ))),
      
      harvest_0_10 = rowSums(across(any_of(
        cols_0_10
      ))),
      
      harvest_10_25 = rowSums(across(any_of(
        cols_10_25
      ))),
      
      harvest_10_20 = rowSums(across(any_of(
        cols_10_20
      ))),
      harvest_20_30 = rowSums(across(any_of(
        cols_20_30
      ))),
      harvest_30_40 = rowSums(across(any_of(
        cols_30_40
      ))),
      harvest_gt_40 = rowSums(across(any_of(
        cols_gt_40
      )))
      
    ) %>%
    
  # move the columns that aren't HFI features or landcover to the front
  select(array, site, buffer_dist, array_year, array_visit,
         harvest_0_15, 
         harvest_gt_15,
         
         harvest_0_10,
         harvest_10_25,
         harvest_gt_25,
         
         harvest_10_20,
         harvest_20_30,
         harvest_30_40,
         harvest_gt_40)
    
  }
  )
```

## 1.4. Add the harvest and do some additional grouping

```{r}
hfi_final <- hfi_grouped %>% 
  
  left_join(hfi_harvest, by = c("array", "site", "array_visit", "buffer_dist", "array_year")) %>%

  mutate(
    # create column osm_industrial following Marissa's convention
    osm_industrial = 
      borrowpits +
      clearings +
      facilities +
      mines,
    
    # Lump popeline and transmission line
    pipe_trans =
      pipeline +
      transmission_lines,
    
    # Lump seismic
    seismic_total = 
      seismic_lines_conventional + 
      seismic_lines_3D,
    
    # Cumulative site disturbance without harvest
    cfi =
      osm_industrial +
      wells_total + 
      seismic_total + 
      pipe_trans +
      trails +
      roads + 
      railways,
    
    # Cumulative site disturbance with veg edges
    cfi_with_vegedges = 
      cfi + 
      veg_edges,
    
        # Cumulative site disturbance with harvest
    cfi_with_harvest = 
      cfi_with_vegedges + 
      harvest_0_10,

    ) %>%
  
  # remove other features we don't need
  select(!any_of(c(
    'cfo',
    'cultivation',
    'reclaimed',
    'recreation',
    'residential',
    'water',
    'borrowpits',
    'clearings',
    'facilities',
    'mines',
    'pipeline',
    'transmission_lines',
    'landfills',
    'other_linear_features',
    'woody_vegetation_removal'
  ))) %>%
  
  relocate(array, site, buffer_dist, array_year, array_visit, sort(names(.)))

names(hfi_final)
```

Looks good!!

# 3. Tidy up the forest inventory data

## 3.1. Clean up and add sample year column

```{r}
sbfi <- sbfi %>%
  
  set_names(tolower(names(.)) %>%
              str_replace_all(., pattern = c('-'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_') %>%  # what you want the pattern to be replaced with
              str_replace_all(., pattern = c('\\.'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_')
            ) %>%

  # What year were the arrays sampled? Used to determine recent fire.
  mutate(array_year = as.numeric(str_extract(array_visit, "\\d+$"))) %>%
  relocate(array_year, .after = array_visit) %>%
  
  mutate(
    # fire
    fire_total = rowSums(across(contains('fire_pct')))
    )

  
names(sbfi)
```

## 3.2. Calculate fire by year columns

```{r}
# Now compute row-wise fire_0_15 and fire_gt_15
# Loop through the sample years. This is the most efficient code I could think of.  
sbfi_fire <- purrr::map_dfr(unique(sbfi$array_year), ~{
  
  # Columns to sum (based on fire year and sample year)
  cols_0_15 <- paste0("fire_pct_", seq(.x - 14, .x))
  cols_gt_15 <- paste0("fire_pct_", seq(1900, .x - 15))
  
  # By decade
  cols_0_10 <- paste0("fire_pct_", seq(.x - 9, .x))
  cols_10_20 <- paste0("fire_pct_", seq(.x - 19, .x - 10))
  cols_20_30 <- paste0("fire_pct_", seq(.x - 29, .x - 20))
  cols_30_40 <- paste0("fire_pct_", seq(.x - 39, .x - 30))
  cols_gt_40 <- paste0("fire_pct_", seq(1900, .x - 40))
  
  # Establishment and aggrandization stages:
  cols_10_25 <- paste0("fire_pct_", seq(.x - 24, .x - 10))
  cols_gt_25 <- paste0("fire_pct_", seq(1900, .x - 25))


  sbfi %>%
    filter(array_year == .x) %>%
    mutate(
      
      # Sum them
      fire_0_15 = rowSums(across(any_of(
        cols_0_15
      ))),
      
      fire_gt_15 = rowSums(across(any_of(
        cols_gt_15
      ))),
      
      fire_gt_25 = rowSums(across(any_of(
        cols_gt_25
      ))),
      
      fire_10_25 = rowSums(across(any_of(
        cols_10_25
      ))),
      
      fire_0_10 = rowSums(across(any_of(
        cols_0_10
      ))),
      fire_10_20 = rowSums(across(any_of(
        cols_10_20
      ))),
      fire_20_30 = rowSums(across(any_of(
        cols_20_30
      ))),
      fire_30_40 = rowSums(across(any_of(
        cols_30_40
      ))),
      fire_gt_40 = rowSums(across(any_of(
        cols_gt_40
      )))
      
    ) %>%
    
  # move the columns that aren't SBFI features or landcover to the front
  select(array, site, buffer_dist, array_year, array_visit,
         fire_0_15, 
         fire_gt_15,
         fire_0_10,
         fire_10_25,
         fire_gt_25,
         fire_10_20,
         fire_20_30,
         fire_30_40,
         fire_gt_40)
  }
  )
```

## 3.3. Finalize and give the species a more meaningful name

```{r}
sbfi_final <- sbfi %>% 
  
  # Some columns we don't want
  select(-contains("fire"),
         -contains("age"), 
         -lc_rock_rubble,
         -lc_snow_ice,
         -lc_bryoids,
         -lc_exposed_barren,
         # We'll get harvest from the ABMI data instead
         -contains("harvest"),
         # We DO want fire total though. Overrides previous removal.
         fire_total) %>%
  
  # Join the fire data
  left_join(sbfi_fire, by = c("array", "site", "buffer_dist", "array_year", "array_visit")) %>%

  relocate(array, site, buffer_dist, array_year, array_visit, sort(names(.))) %>%
  
  # Nicer species column names
  rename_with(
    ~ paste0("pct_", str_remove(., "_pct_of_treed")),
    .cols = contains("_pct_of_treed")
  )

names(sbfi_final)
```


# 4. Format the fire covs from Alberta wildfire service

## 4.1. Initial cleaning, add the sample year column

```{r}
fire <- fire %>%
  
  set_names(tolower(names(.)) %>%
              str_replace_all(., pattern = c('-'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_') %>%  # what you want the pattern to be replaced with
              str_replace_all(., pattern = c('\\.'), # provide the character pattern to look for (if you don't keep the \\ it won't work)
                              replacement = '_')
            ) %>%

  # What year were the arrays sampled? Used to determine recent fire and burns.
  mutate(array_year = as.numeric(str_extract(array_visit, "\\d+$"))) %>%
  relocate(array_year, .after=array_visit)
```

```{r}
names(fire)
```

## 4.2. Calculate fire by year columns for each burn class

This is quite clunky, but basically for each row (with a respective sample year), it identifies which columns are within the time range (using some character parsing). Since we have so many burn codes
```{r}
# Now compute row-wise fire_0_15 and fire_gt_15
# Loop through the sample years FOR EACH BURN CODE. This is the most efficient code I could think of.  
burn_class_strings <- c("fire_class1_",
                       "fire_class2_",
                       "fire_class3_",
                       "fire_class4_",
                       "fire_class5_",
                       "fire_class9_"
                       )

# Loop through burn classes
fire_binned <- purrr::map(burn_class_strings, function(burn_class) {
  
  # Loop through the array years
  purrr::map_dfr(unique(fire$array_year), function(year) {
    
    # Columns to sum (based on fire year and sample year)
    cols_0_15 <- paste0(burn_class, seq(year - 14, year))
    cols_gt_15 <- paste0(burn_class, seq(1900, year - 15))
    
    # By decade
    cols_0_10 <- paste0(burn_class, seq(year - 9, year))
    cols_10_20 <- paste0(burn_class, seq(year - 19, year - 10))
    cols_20_30 <- paste0(burn_class, seq(year - 29, year - 20))
    cols_30_40 <- paste0(burn_class, seq(year - 39, year - 30))
    cols_gt_40 <- paste0(burn_class, seq(1900, year - 40))
    
    # Establishment and aggrandization stages:
    cols_10_25 <- paste0(burn_class, seq(year - 24, year - 10))
    cols_gt_25 <- paste0(burn_class, seq(1900, year - 25))
    
    # Total fire across all possible years!
    cols_total <- paste0(burn_class, seq(1900, 2030))
  
  
    fire %>%
      filter(array_year == year) %>%
      
      mutate(
        
        # Use glue syntax "{burn_class}" and the walrus operator :=
        "{burn_class}0_15" := rowSums(across(any_of(cols_0_15))),
        "{burn_class}gt_15" := rowSums(across(any_of(cols_gt_15))),
        
        "{burn_class}gt_25" := rowSums(across(any_of(cols_gt_25))),
        "{burn_class}10_25" := rowSums(across(any_of(cols_10_25))),
        
        "{burn_class}0_10" := rowSums(across(any_of(cols_0_10))),
        "{burn_class}10_20" := rowSums(across(any_of(cols_10_20))),
        "{burn_class}20_30" := rowSums(across(any_of(cols_20_30))),
        "{burn_class}30_40" := rowSums(across(any_of(cols_30_40))),
        "{burn_class}gt_40" := rowSums(across(any_of(cols_gt_40))),
        
        "{burn_class}total" := rowSums(across(any_of(cols_total))),

        
      ) %>%
      
    # Drop the original single-year columns
    select(-any_of(c(outer(burn_class_strings, 1900:2030, paste0)))) %>%
    
    # move the columns that aren't SBFI features or landcover to the front
    relocate(array, site, buffer_dist, array_year, array_visit)
      
    # Close year loop
    })
  
  # Close burn class loop
  }) %>% 
  
  # Name the resulting list elements based on the burn classes
  purrr::set_names(burn_class_strings) %>%
  
  # Apply left_join across the whole list. We could use cbind but then the ID columns would be duplicated. 
  reduce(left_join, by = c("array", "site", "buffer_dist", "array_year", "array_visit")) 

str(fire_binned)
```


## 4.3. Finish formatting the fire data

Let's sum across all burn codes for the time bins. 
This is probably what we'll use for modelling anyway due to data density.
Clunky BUT IT WORKS AND IT'S VERY EASY TO READ. 

```{r}
bins <- c("0_15", "gt_15", "0_10", "10_25", "gt_25", "10_20", "20_30", "30_40", "gt_40", "total")

fire_final <- fire_binned %>%
  
  mutate(
    # 0 to 15 years
    fire_0_15 = rowSums(across(contains("_0_15")), na.rm = TRUE),
    
    # Greater than 15 years
    fire_gt_15 = rowSums(across(contains("_gt_15")), na.rm = TRUE),
    
    # 0 to 10 years
    fire_0_10 = rowSums(across(contains("_0_10")), na.rm = TRUE),
    
    # 10 to 25 years
    fire_10_25 = rowSums(across(contains("_10_25")), na.rm = TRUE),
    
    #Greater than 25 years
    fire_gt_25 = rowSums(across(contains("_gt_25")), na.rm = TRUE),
    
    # 10 to 20 years
    fire_10_20 = rowSums(across(contains("_10_20")), na.rm = TRUE),
    
    # 20 to 30 years
    fire_20_30 = rowSums(across(contains("_20_30")), na.rm = TRUE),
    
    # 30 to 40 years
    fire_30_40 = rowSums(across(contains("_30_40")), na.rm = TRUE),
    
    # Greater than 40 years
    fire_gt_40 = rowSums(across(contains("_gt_40")), na.rm = TRUE),
    
    # Total fire footprint across all classes
    fire_total = rowSums(across(contains("_total")), na.rm = TRUE)
  )

```


# 5. Put it all together:

## 5.1. Import lat/long and UTM 12N coordinates

```{r}
# read in the 'site coordinate' files. Purrr works for now. 
coords <- read_csv("./data/input/OSM_site_coordinates_2021-2024.csv")
``` 

## 5.2. Remove some columns for now...

First SBFI fire. We don't want these for now, since we have better fire polygons from AB wildfire now!!

```{r}
names(sbfi_final)

# Remove all fire cols
sbfi_final <- sbfi_final %>%
  
  select(-contains("fire"))

names(sbfi_final)
```

Second let's remove the burn_class columns until we have a specific question that will use them, just to keep the dataframe more readable. 

```{r}
names(fire_final)

# Remove all fire cols
fire_final <- fire_final %>%
  
  select(-contains("fire_class"))

names(fire_final)
```

## 5.3. Append everything and final inspection

```{r}
# Tidy up and final check
covs <- sbfi_final %>%
  
  left_join(hfi_final, by = c("array", "site", "buffer_dist", "array_visit", "array_year")) %>%
  
  left_join(fire_final, by = c("array", "site", "buffer_dist", "array_visit", "array_year")) %>%
  
  left_join(ndvi, by = c("array", "site", "buffer_dist", "array_visit")) %>%
  
  left_join(daymet, by = c("array", "site", "array_visit")) %>%

  left_join(coords, by = c("array", "site", "array_visit")) %>%
  
  relocate(array, site, array_year, array_visit, lat, long, easting_12n, northing_12n, start_date, end_date, camera_failure_details, deployment_id, buffer_dist)

summary(covs)
head(covs)
names(covs)

# This should be an even 26 (or whatever number of buffers you've extracted)
nrow(covs) / nrow(coords)
```

## 5.4. Save the data:

```{r}
write_csv(covs, "./data/output/OSM_site_covariates_2021-2024.csv")
```



End script




