---
title: "Visualize OSM mammal data"
author: "Marissa Dyck, modified by Aidan Brushett"
date: "2025-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Before you begin

## Notes

A few notes about this script.

If you are running this with the 2024-2025 data make sure you download the whole (OSM_2024-2025 GitHub repository)[https://github.com/ACMElabUvic/OSM_2024-2025] from the ACMElabUvic GitHub. This will ensure you have all the files, data, and proper folder structure you will need to run this code and associated analyses.

Also make sure you open RStudio through the R project (OSM_2024-2025.Rproj) this will automatically set your working directory to the correct place (wherever you saved the repository) and ensure you don't have to change the file paths for some of the data. 

If you have question please email the most recent author, currently   

Marissa A. Dyck   
Postdoctoral research fellow    
University of Victoria    
School of Environmental Studies     
Email: [marissadyck17@gmail.com](marissadyck17@gmail.com)     

**and**

Aidan M. Brushett  
M.Sc. student
University of Victoria    
School of Environmental Studies     
Email: [aidanbrushett@uvic.ca](aidanbrushett@uvic.ca)     


## R and RStudio

Before starting you should ensure you have the latest version of R and RStudio downloaded. This code was generated under R version 4.2.3 and with RStudio version 2024.04.2+764.    

You can download R and RStudio [HERE](https://posit.co/download/rstudio-desktop/)   


## R markdown

This script is written in R markdown and thus uses a mix of coding markup languages and R. If you are planning to run this script with new data or make any modifications you will want to be familiar with some basics of R markdown.

Below is an R markdown cheatsheet to help you get started,    
[R markdown cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)    

# 0. Setup

## Install packages

If you don't already have the following packages installed, use the code below to install them. *NOTE this will not run automatically as eval=FALSE is included in the chunk setup (i.e. I don't want it to run every time I run this code since I have the packages installed)

```{r install packages, eval=FALSE}

install.packages('tidyverse') 
install.packages('withr') 
```

## Load libraries

Then load the packages to your library.

```{r libraries}
rm(list=ls()) # start with a fresh global environment
library(tidyverse) # data tidying, visualization, and much more; this will load all tidyverse packages, can see complete list using tidyverse_packages()
```

> Note: all the processing of response data was performed in script 1, and landscape data in script 2. We will use the products in ./data/processed/ for this script. 

# 1. Response Variable figures (MODIFY FOR FYE DEADLINE AS NEEDED)

## 1.1. Import data

Let's read in the data from script 3. Importing the *entire* dataset means we can reuse this code for the data paper, or project-wide summaries as needed without major revision. 

```{r}
# Operability adjusted for snow cover
operating <- read_csv("./data/processed/OSM_operating_days_2021-2024.csv")
```


```{r}
# Independent detections
indet <- read_csv("./data/processed/OSM_independent_detections_2021-2024.csv")
```

```{r}
# Total detections and presence per site:
totaldet <- read_csv("./data/processed/OSM_response_total_detections_per_effort_2021-2024.csv")
```


```{r}
# Landscape covariates
covariates <- read_csv("./data/processed/OSM_site_covariates_2021-2024.csv")
```


```{r}
# Landscape unit classifications. 
lu_classification <- read_csv("./data/raw/landscape_unit_land_use_classifications.csv")
```

What sites actually have camera data? We will ONLY use these sites for FYE figures. 
This handles missing data from image tagging that we should track down later, but maybe after FYE. 

```{r}
sites_to_plot <- (unique(indet$site))
```



# 2. Plot camera operability

Let's read in the camera operability dates *for cameras with data* and plot it out. For now, we'll focus this entire script on just the 2024-2025 data since that's what we want for the Fiscal report. Importing the *entire* dataset means we can reuse this code for the data paper, or project-wide summaries as needed without major revision. 

Now we can plot it:


First, the complete OSM operating history:
```{r}
operating %>%
  
  # Which days were the arrays working? 
  # We want to collapse the site-by-site data here for a cleaner graph
  distinct(array, date) %>%
  
  # Add in the classification column
  left_join(lu_classification, by = "array") %>%
  
  # Set the order for the legend in increasing disturbance
  mutate(classification = factor(classification, 
                                 levels = c("low development", "pre in situ", "in situ", "mine"))) %>%
  
  # Make a basic Gantt chart colored by classification
  ggplot(., aes(x = date, y = array, fill = classification)) +
    geom_tile(height = 0.8) +
    scale_x_date(date_breaks = "6 months", date_labels = "%b '%y") +
    scale_fill_viridis_d() +
    theme_classic() + 
    labs(x = "Date",
         y = "Landscape Unit",
         fill = "Landscape \nClassification")
```

Save it: 

```{r}
ggsave("./figures/OSM_operability_allyears.png", width = 7, height = 4)
```

Second, just the 2024-2025 OSM operating history:

```{r}
operating %>%  
  
  # This query should pull out only the 2024-2025 data from these arrays :)
  filter(year >= 2024,
         array %in% c("LU4", "LU2", "LU3", "LU8")) %>%

  # Filter to sites that are *not* missing data
  filter(site %in% sites_to_plot) %>%
  
  # Basic tile chart. Colored by array to keep things clear 
  ggplot(., aes(x = date, y = site, fill = array)) +
    geom_tile() +
    scale_x_date(date_breaks = "2 months", date_labels = "%b '%y") +
    theme_classic() + 
    scale_fill_viridis_d() +
    theme(axis.text.y = element_blank()) +
    labs(x = "Date",
         y = "Camera Station",
         fill = "Landscape \nUnit")
```

Save it: 

```{r}
ggsave("./figures/OSM_operability_2024-2025.png", width = 7, height = 4)
```



# 3. Independent detections and naive occupancy

## 3.1. Identify focal species

We'll identify some focal species (not used for now) and also bin some of the similar species categories. 

```{r}
# We will not use this for now
focal_species <- tolower(c('Red squirrel',
                   'Snowshoe hare', 
                   'White-tailed deer',
                   'Black bear',
                   'Coyote',
                   'Moose',
                   'Lynx', 
                   'Unknown',
                   'Marten',
                   'Wolverine',
                   'Grey wolf',
                   'Caribou',
                   'Red fox', 
                   'Fisher',
                   'Spruce grouse',
                   'Ruffed grouse', 
                   'Grey jay',
                   'Long-tailed weasel',
                   'Ruffed grouse',
                   'Short-tailed weasel',
                   'Otter',
                   'Canada goose',
                   'Beaver',
                   'Human (various)',
                   'Porcupine',
                   'Owl',
                   'Cougar'
                   ))

human_species <- tolower(c('ATVer', 
                   'Human',
                   'Hunter',
                   'Hiker',
                   'Snowmobiler'
                   ))

unknown_species <- tolower(c('Unknown',
                     'Unknown canid', 
                     'Unknown bird',
                     'Unknown ungulate',
                     'Unknown deer',
                     'Unknown mustelid',
                     'Unknown bear'))
```

## 3.2. Independent detections and naive occupancy per array

```{r}
# Split the data up by each visit to an array 
# i.e., we do this separate for repeat samples of the same LU)
for(LU in unique(totaldet$array_visit)){
  
  # Track progress in loop
  print(LU)

  # Create the dataset for the figure
  fig_totaldet_data <- totaldet %>%
    
    # Only use sites with data
    filter(site %in% sites_to_plot) %>%
    
    # Filter to a specific sample of an array
    filter(array_visit == LU) %>%
    
    # Group together a few similar species
    mutate(species = case_when(
      # Lump all human-use together
      species %in% human_species ~ "human (various)",
      # Lump most unknown together
      species %in% unknown_species ~ "unknown (various)",
      TRUE ~ species)) %>%
    
    # Remove the species we don't want in the figure
    filter(!species %in% c("staff", 
                           "other", 
                           "other birds")) %>%
    
    # Add in the classification column
    left_join(lu_classification, by = "array") %>%

    group_by(species, array, classification) %>%
    
    # How many rows (detections) per species? And other metrics
    summarize(detections = sum(detections),
              effort = sum(camera_days),
              present_sites = sum(presence), 
              total_sites = n()) %>%
    
    # Calculate response metric (naive occupancy)
    mutate(naiveocc = present_sites / total_sites)
  
  
  # Create the title for the figure separately here
  # Array name and LU classification
  title <- paste0(unique(fig_totaldet_data$array),
                  " (",
                  unique(fig_totaldet_data$classification),
                  ")")
  
  
  # Create the total detections figure
  fig_totaldet_data %>%
    
    # Basic bar graph with a nice title
    ggplot(aes(x = detections, y = fct_reorder(species, detections))) +
    geom_col(fill = "gray30") +
    geom_text(aes(label = detections), hjust = -0.2, size = 3) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
    theme_classic() +
    labs(
      x = "Total detections",
      y = NULL,
      title = title
    )
  
  # Save the figure
  ggsave(paste0("./figures/arrays/", LU, "_independent_detections.png"), width = 8, height = 5, dpi = 500) 

  
    # Create the total detections figure
  fig_totaldet_data %>%
    ggplot(aes(x = naiveocc, y = fct_reorder(species, naiveocc))) +
    geom_col(fill = "gray30") +
    geom_text(aes(label = round(naiveocc, 2)), hjust = -0.2, size = 3) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
    theme_classic() +
    labs(
      x = "Naive occupancy",
      y = NULL,
      title = title
    )

  # Save the figure
  ggsave(paste0("./figures/arrays/", LU, "_naiveoccupancy.png"), width = 8, height = 5, dpi = 500) 
}
# end for loop
```



# 4. Independent detections per repeat sample:

```{r}
# Iterate through each LU to generate array-wise comparisons
for(LU in c("LU2", "LU3")){
  
  # Data Wrangling: Filter and aggregate detection data
  fig_paireddet_data <- totaldet %>% 
    
    # Isolate the current LU across all years
    filter(array == LU) %>%
    filter(site %in% sites_to_plot) %>%
    
    # Aggregate miscellaneous species into broader categories
    mutate(species = case_when(
        species %in% human_species ~ "human (various)",
        species %in% unknown_species ~ "unknown (various)",
        TRUE ~ species)) %>%
      
    # Remove non-target species or human-activity markers
    filter(!species %in% c("staff", 
                             "other", 
                             "other birds",
                             "human (various)",
                             "unknown (various)")) %>%
    
    # Group by survey year and species to calculate summary statistics
    group_by(array, array_visit, species) %>%
    
    summarize(detections = sum(detections),
              camera_days = sum(camera_days),
              present_sites = sum(presence), 
              total_sites = n(),
              .groups = "drop") %>% # Ungroup to prevent downstream errors
    
    # Calculate derived ecological metrics
    mutate(naiveocc = present_sites / total_sites, # Proportion of sites occupied
           detections_per_effort = detections/camera_days * 100, # Detection rate per 100 days
           # Extract year from the visit ID (e.g., "LU2_2021" becomes "2021")
           array_visit = str_replace(array_visit, paste0(LU, "_"), ""))
  
  
  ###################
  # PLOT 1: Adjusted for Effort
  # Standardizing detections by camera-days accounts for variation in sampling duration
  ggplot(fig_paireddet_data, aes(x = detections_per_effort, 
                  y = species, 
                  fill = factor(array_visit, levels = c(2024, 2021)))) +
      # Basic bar chart
      geom_bar(stat = "identity", position = "dodge") +
    
      scale_fill_manual(values = c("grey60", "grey20"),
                        breaks = c("2021", "2024")) +
    
      scale_x_continuous(expand = expansion(mult = c(0, 0.05))) + 
  
      theme_classic() + 
      # Title should have LU name and land-use classification appended together    
      labs(y = "",
           x = "Detections per 100 sampling days",
           title = paste0(LU, ' (', 
                          subset(lu_classification, array == LU)$classification,
                          ')'),
           fill = "Survey Year")
  
  # Save plot
  ggsave(paste0("./figures/arrays/", LU, "_detections_adjusted_2021_2024.png"), width = 7, height = 4)

  
  ###################
  # PLOT 2: Raw Detections (Not adjusted for effort)
  # Visualizes total volume of detections recorded per species
  ggplot(fig_paireddet_data, aes(x = detections, 
                  y = species, 
                  fill = factor(array_visit, levels = c(2024, 2021)))) +
      # Basic bar chart
      geom_bar(stat = "identity", position = "dodge") +
    
      scale_fill_manual(values = c("grey60", "grey20"),
                        breaks = c("2021", "2024")) +
    
      scale_x_continuous(expand = expansion(mult = c(0, 0.05))) + 
  
      theme_classic() + 
      # Title should have LU name and land-use classification appended together    
      labs(y = "",
           x = "Total detections",
           title = paste0(LU, ' (', 
                          subset(lu_classification, array == LU)$classification,
                          ')'),
           fill = "Survey Year")

  # Save plot
  ggsave(paste0("./figures/arrays/", LU, "_detections_2021_2024.png"), width = 7, height = 4)
  
  
  ###################
  # PLOT 3: Naive Occupancy
  # Visualizes spatial distribution (presence/absence) across the array
  ggplot(fig_paireddet_data, aes(x = naiveocc, 
                  y = species, 
                  fill = factor(array_visit, levels = c(2024, 2021)))) +
      # Basic bar chart
      geom_bar(stat = "identity", position = "dodge") +
    
      scale_fill_manual(values = c("grey60", "grey20"),
                        breaks = c("2021", "2024")) +
    
      scale_x_continuous(expand = expansion(mult = c(0, 0.05))) + 
  
      theme_classic() + 
    
      # Title should have LU name and land-use classification appended together
      labs(y = "",
           x = "Naive occupancy",
           title = paste0(LU, ' (', 
                          subset(lu_classification, array == LU)$classification,
                          ')'),
           fill = "Survey Year")

  # Save plot
  ggsave(paste0("./figures/arrays/", LU, "_naiveoccupancy_2021_2024.png"), width = 7, height = 4)
}
```



# 5. Other mammal figures

A few quick figures to look at the detection data in other ways

First create a vector of the list of mammals to use for quick data visualization/exploration.

```{r graph mammal detections}
mammals <- c('White-tailed deer',
             'Black bear',
             'Snowshoe hare',
             'Moose',
             'Coyote',
             'Fisher',
             'Red squirrel',
             'Striped skunk',
             'Grey wolf',
             'Red fox',
             'Cougar',
             'Lynx',
             'Short-tailed weasel',
             'Porcupine',
             'Beaver',
             'Martin',
             'Wolverine',
             'Caribou',
             'Long-tailed weasel')
```

## 5.1. Total independent detections for all species, combined across all 2024-2025 sites

Plot it: 

```{r graph detections}
# remove NAs and select just images with mammals first then pipe new data into ggplot
indet %>% 
  
    # This query should pull out only the 2024-2025 data from these arrays :)
  filter(year >= 2024,
         array %in% c("LU4", "LU2", "LU3", "LU8")) %>%

  # retain only the species in `mammals` that we want to plot:
  #filter(species %in% tolower(mammals)) %>% 
  
  # get the number of individual detections per species to add to graph
  group_by(species) %>% 
  
  # How many rows (detections) are there for a species, *combined across all sites*??
  mutate(n = n()) %>% 
  
  ungroup() %>% 
  
  # Create
  ggplot(., aes(x = species)) +
    # create bar graph of the counts of each spp in the data
    geom_bar() +
    # add the number of detections above each bar using the variable n we calculated earlier
    geom_text(aes(label = n,
                  y = n + 100),
              size = 3) +
    # change y axis label
    labs(y = 'Detections',
         x = "", 
         title = "Total detections for 2024-2025 sampling year (LU2, LU3, LU8)") +
    # change breaks for y axis
    scale_y_continuous(breaks = seq(0,3500, by = 250),
                       expand = expansion(mult = c(0, 0.05))) +
    theme_classic() + 
    # change theme elements
    theme(axis.text.x = element_text(angle = 45,
                                     hjust = 1,
                                     vjust = 1),
          panel.grid = element_blank())
```

If we want to save the plot for easier viewing later we can use the code below:

```{r save det_graph}

# save graph as jpeg (can also save as tiff, png, pdf by changing the file extension) but don't use .tiff in the github repo it takes up too much space and causes issues
ggsave('./figures/total_detections_2024-2025.png',
       width = 8,
       height = 5,
       units = 'in',
       dpi = 600)

```


## 5.2. TOTAL stats per LU. 

```{r}
# Create the dataset for the figure
fig_totaldet_data <- totaldet %>%
    
    # Only sites for which there is data (remove missing sites)
    filter(site %in% sites_to_plot) %>%
    
    # Group together a few similar species
    mutate(species = case_when(
      species %in% human_species ~ "human (various)",
      species %in% unknown_species ~ "unknown (various)",
      TRUE ~ species)) %>%
    
    # Remove the species we don't want in the figure
    filter(!species %in% c("staff", 
                           "other", 
                           "other birds")) %>%
    
   # Add the 'classification' column back in
    left_join(lu_classification, by = "array") %>%

    group_by(species, array, array_visit, classification) %>%
    
    # How many rows (detections) per species?
    summarize(detections = sum(detections),
              effort = sum(camera_days),
              present_sites = sum(presence), 
              total_sites = n()) %>%
    
    # Calculate the adjusted metrics that we want to present
    mutate(naiveocc = present_sites / total_sites,
           detections_per_effort = detections/effort*100) %>%
  
    ungroup()



# DETECTIONS PER EFFORT FOR EACH ARRAY
fig_totaldet_data %>%
  
  # Set Classification order
  mutate(classification = factor(classification, 
                                 levels = c("low development", "pre in situ", "in situ", "mine"))) %>%
  
  # Arrange by classification to group LUs together on the x-axis
  arrange(classification, array_visit) %>%
  
  # Lock in the order of array_visit based on the arrange step above
  mutate(array_visit = factor(array_visit, levels = unique(array_visit))) %>%  
  
  # Plot a basic bar chart, with separate panels per species
  ggplot(aes(x = array_visit, y = detections_per_effort, fill = classification)) +
    geom_col() +
    facet_wrap(~ species, scales = "free_y") +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
      strip.background = element_blank(),
      panel.grid = element_blank()
    ) + 
    labs(x = "Landscape Unit",
         y = "Detections per 100 sampling days",
         fill = "Landscape \nClassification")

# Save the figure
ggsave("./figures/total_detections_per_array.png", width = 18, height = 11)





# NAIVE OCCUPANCY FOR EACH ARRAY
fig_totaldet_data %>%
  
  # Set Classification order
  mutate(classification = factor(classification, 
                                 levels = c("low development", "pre in situ", "in situ", "mine"))) %>%
  
  # Arrange by classification to group LUs together on the x-axis
  arrange(classification, array_visit) %>%
  
  # Lock in the order of array_visit based on the arrange step above
  mutate(array_visit = factor(array_visit, levels = unique(array_visit))) %>%  
  
  # Plot a basic bar chart, with separate panels per species
  ggplot(aes(x = array_visit, y = naiveocc, fill = classification)) +
    geom_col() +
    facet_wrap(~ species, scales = "free_y") +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
      strip.background = element_blank(),
      panel.grid = element_blank()
    ) + 
    labs(x = "Landscape Unit",
         y = "Naive occupancy",
         fill = "Landscape \nClassification")

# Save the figure
ggsave("./figures/naiveoccupancy_per_array.png", width = 18, height = 11)
```










# 2. Covariate figures

## 2.1. Covariate histogram by LU

First, make the data frame that we'll use:

```{r}
fig_covariate_data <- covariates %>%
  
  # Add the 'in situ' etc. classification
  left_join(lu_classification, by = "array") %>%
  mutate(classification = factor(classification, 
                                 levels = c("low development", "pre in situ", "in situ", "mine"))) %>%
  
  # Remove the early, partial sample of LU2 and LU3 (the only arrays from 2022)
  filter(array_year >= 2022,
         # Buffer 500m is the geometric mean of our sampled scales, let's use it. 
         buffer_dist == 500)
```

Start with seismic lines. Could a density plot look good?

```{r}
fig_covariate_data %>%

# Instead of a histogram, what if we plotted density transparently on the same figure??  
ggplot(aes(x = seismic_total, group = classification, fill = classification)) +
    geom_density(alpha = 0.4, color = "white", size = 0.2) + # Transparent overlapping blobs
    scale_fill_viridis_d() +
    theme_classic() +
    labs(x = "Seismic Line Footprint (%)",
         y = "Density",
         fill = "Landscape \nClassification")
```

This was a cool idea but kind of ugly. Back to histograms:

```{r}
fig_covariate_data %>%
  
    
  # 1. Rank each array within its own classification group
  group_by(classification) %>%
  mutate(group_rank = as.numeric(factor(array))) %>%
  ungroup() %>%
  
  # Arrange by rank first, then by classification
  # This creates the order: Rank1-Low, Rank1-Pre, Rank1-InSitu, Rank1-Mine, Rank2-Low...
  arrange(group_rank, classification) %>%
  
  # 3. Lock in this specific sequence for the facets
  mutate(array = fct_inorder(array)) %>%
  
# Basic histogram code, 15 bins looks nice
ggplot(aes(x = seismic_total, fill = classification)) +
    geom_histogram(bins = 15) +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +

    # Facet by classification to separate the distributions
    facet_wrap(~ array) +
    
    theme_bw() +
    theme(strip.background = element_blank(),
          panel.grid = element_blank()) +
    labs(x = "Seismic line footprint (%)",
         y = "Number of sites",
         fill = "Landscape \nClassification")

ggsave("./figures/covariates/seismic_histogram_all_arrays.png", width = 10, height = 5)
```

Well pads:

```{r}
fig_covariate_data %>%

    
  # 1. Rank each array within its own classification group
  group_by(classification) %>%
  mutate(group_rank = as.numeric(factor(array))) %>%
  ungroup() %>%
  
  # 2. Arrange by rank first, then by classification
  # This creates the order: Rank1-Low, Rank1-Pre, Rank1-InSitu, Rank1-Mine, Rank2-Low...
  arrange(group_rank, classification) %>%
  
  # 3. Lock in this specific sequence for the facets
  mutate(array = fct_inorder(array)) %>%
  
  
# Basic histogram code, 15 bins looks nice  
ggplot(aes(x = wells_total, fill = classification)) +
    geom_histogram(bins = 15) +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +

    # Facet by classification to separate the distributions
    facet_wrap(~ array) +
    
    theme_bw() +
    theme(strip.background = element_blank(),
          panel.grid = element_blank()) +
    labs(x = "Well pad footprint",
         y = "Number of sites",
         fill = "Landscape \nClassification")

ggsave("./figures/covariates/wellpad_histogram_all_arrays.png", width = 10, height = 5)

```

Harvest areas:

```{r}
fig_covariate_data %>%
  
  
  # 1. Rank each array within its own classification group
  group_by(classification) %>%
  mutate(group_rank = as.numeric(factor(array))) %>%
  ungroup() %>%
  
  # 2. Arrange by rank first, then by classification
  # This creates the order: Rank1-Low, Rank1-Pre, Rank1-InSitu, Rank1-Mine, Rank2-Low...
  arrange(group_rank, classification) %>%
  
  # 3. Lock in this specific sequence for the facets
  mutate(array = fct_inorder(array)) %>%
  
  
# Basic histogram code, 15 bins looks nice
ggplot(aes(x = harvest_total, fill = classification)) +
    geom_histogram(bins = 15) +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +

    # Facet by classification to separate the distributions
    facet_wrap(~ array) +
    
    theme_bw() +
    theme(strip.background = element_blank(),
          panel.grid = element_blank()) +
    labs(x = "Timber harvest footprint",
         y = "Number of sites",
         fill = "Landscape \nClassification")

ggsave("./figures/covariates/timberharvest_histogram_all_arrays.png", width = 10, height = 5)

```

Industrial facilities: 

```{r}
fig_covariate_data %>%
  
    
  # 1. Rank each array within its own classification group
  group_by(classification) %>%
  mutate(group_rank = as.numeric(factor(array))) %>%
  ungroup() %>%
  
  # 2. Arrange by rank first, then by classification
  # This creates the order: Rank1-Low, Rank1-Pre, Rank1-InSitu, Rank1-Mine, Rank2-Low...
  arrange(group_rank, classification) %>%
  
  # 3. Lock in this specific sequence for the facets
  mutate(array = fct_inorder(array)) %>%
  
# Basic histogram code, 15 bins looks nice
ggplot(aes(x = osm_industrial, fill = classification)) +
    geom_histogram(bins = 15) +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +

    # Facet by classification to separate the distributions
    facet_wrap(~ array, ncol = 4) +
    
    theme_bw() +
    theme(strip.background = element_blank(),
          panel.grid = element_blank()) +
    labs(x = "Industrial processing facilities",
         y = "Number of sites",
         fill = "Landscape \nClassification")

ggsave("./figures/covariates/osmindustrial_histogram_all_arrays.png", width = 10, height = 5)

```


Cumulative footprint:

```{r}
fig_covariate_data %>%
  
    
  # 1. Rank each array within its own classification group
  group_by(classification) %>%
  mutate(group_rank = as.numeric(factor(array))) %>%
  ungroup() %>%
  
  # 2. Arrange by rank first, then by classification
  # This creates the order: Rank1-Low, Rank1-Pre, Rank1-InSitu, Rank1-Mine, Rank2-Low...
  arrange(group_rank, classification) %>%
  
  # 3. Lock in this specific sequence for the facets
  mutate(array = fct_inorder(array)) %>%
  
# Basic histogram code, 15 bins looks nice
ggplot(aes(x = cfi_with_vegedges, fill = classification)) +
    geom_histogram(bins = 15) +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +

    # Facet by classification to separate the distributions
    facet_wrap(~ array, ncol = 4) +
    
    theme_bw() +
    theme(strip.background = element_blank(),
          panel.grid = element_blank()) +
    labs(x = "Cumulative industrial footprint",
         y = "Number of sites",
         fill = "Landscape \nClassification")

ggsave("./figures/covariates/cfi_histogram_all_arrays.png", width = 10, height = 5)

```



Pipelines and transmission lines:

```{r}
fig_covariate_data %>%
  
    
  # 1. Rank each array within its own classification group
  group_by(classification) %>%
  mutate(group_rank = as.numeric(factor(array))) %>%
  ungroup() %>%
  
  # 2. Arrange by rank first, then by classification
  # This creates the order: Rank1-Low, Rank1-Pre, Rank1-InSitu, Rank1-Mine, Rank2-Low...
  arrange(group_rank, classification) %>%
  
  # 3. Lock in this specific sequence for the facets
  mutate(array = fct_inorder(array)) %>%
  
# Basic histogram code, 15 bins looks nice
ggplot(aes(x = pipe_trans, fill = classification)) +
    geom_histogram(bins = 15) +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +

    # Facet by classification to separate the distributions
    facet_wrap(~ array, ncol = 4) +
    
    theme_bw() +
    theme(strip.background = element_blank(),
          panel.grid = element_blank()) +
    labs(x = "Pipelines & transmission lines",
         y = "Number of sites",
         fill = "Landscape \nClassification")

ggsave("./figures/covariates/pipeline_histogram_all_arrays.png", width = 10, height = 5)

```





Roads:

```{r}
fig_covariate_data %>%
  
    
  # 1. Rank each array within its own classification group
  group_by(classification) %>%
  mutate(group_rank = as.numeric(factor(array))) %>%
  ungroup() %>%
  
  # 2. Arrange by rank first, then by classification
  # This creates the order: Rank1-Low, Rank1-Pre, Rank1-InSitu, Rank1-Mine, Rank2-Low...
  arrange(group_rank, classification) %>%
  
  # 3. Lock in this specific sequence for the facets
  mutate(array = fct_inorder(array)) %>%
  
# Basic histogram code, 15 bins looks nice
ggplot(aes(x = roads, fill = classification)) +
    geom_histogram(bins = 15) +
    scale_fill_viridis_d() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +

    # Facet by classification to separate the distributions
    facet_wrap(~ array, ncol = 4) +
    
    theme_bw() +
    theme(strip.background = element_blank(),
          panel.grid = element_blank()) +
    labs(x = "Roads",
         y = "Number of sites",
         fill = "Landscape \nClassification")

ggsave("./figures/covariates/road_histogram_all_arrays.png", width = 10, height = 5)
```