---
title: "Visualize OSM mammal data"
author: "Marissa Dyck, modified by Aidan Brushett"
date: "2025-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Before you begin

## Notes

A few notes about this script.

If you are running this with the 2024-2025 data make sure you download the whole (OSM_2024-2025 GitHub repository)[https://github.com/ACMElabUvic/OSM_2024-2025] from the ACMElabUvic GitHub. This will ensure you have all the files, data, and proper folder structure you will need to run this code and associated analyses.

Also make sure you open RStudio through the R project (OSM_2024-2025.Rproj) this will automatically set your working directory to the correct place (wherever you saved the repository) and ensure you don't have to change the file paths for some of the data. 

If you have question please email the most recent author, currently   

Marissa A. Dyck   
Postdoctoral research fellow    
University of Victoria    
School of Environmental Studies     
Email: [marissadyck17@gmail.com](marissadyck17@gmail.com)     

**and**

Aidan M. Brushett  
M.Sc. student
University of Victoria    
School of Environmental Studies     
Email: [aidanbrushett@uvic.ca](aidanbrushett@uvic.ca)     


## R and RStudio

Before starting you should ensure you have the latest version of R and RStudio downloaded. This code was generated under R version 4.2.3 and with RStudio version 2024.04.2+764.    

You can download R and RStudio [HERE](https://posit.co/download/rstudio-desktop/)   


## R markdown

This script is written in R markdown and thus uses a mix of coding markup languages and R. If you are planning to run this script with new data or make any modifications you will want to be familiar with some basics of R markdown.

Below is an R markdown cheatsheet to help you get started,    
[R markdown cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)    

# 0. Setup

## Install packages

If you don't already have the following packages installed, use the code below to install them. *NOTE this will not run automatically as eval=FALSE is included in the chunk setup (i.e. I don't want it to run every time I run this code since I have the packages installed)

```{r install packages, eval=FALSE}

install.packages('tidyverse') 
install.packages('withr') 
```

## Load libraries

Then load the packages to your library.

```{r libraries}
rm(list=ls()) # start with a fresh global environment
library(tidyverse) # data tidying, visualization, and much more; this will load all tidyverse packages, can see complete list using tidyverse_packages()
```

> Note: all the processing of response data was performed in script 1, and landscape data in script 2. We will use the products in ./data/processed/ for this script. 

# 1. Response Variable figures (MODIFY FOR FYE DEADLINE AS NEEDED)

## 1.1. Import data

Let's read in the data from script 3. Importing the *entire* dataset means we can reuse this code for the data paper, or project-wide summaries as needed without major revision. 

```{r}
# Operability adjusted for snow cover
operating <- read_csv("./data/processed/OSM_operating_days_2021-2024.csv")
```


```{r}
# Independent detections
indet <- read_csv("./data/processed/OSM_independent_detections_2021-2024.csv")
```


```{r}
# Monthly presence and absence
monthlyprop <- read_csv("./data/processed/OSM_response_monthly_proportional_presence_2021-2024.csv")
```


```{r}
# Total detections and presence per site:
totaldet <- read_csv("./data/processed/OSM_response_total_detections_per_effort_2021-2024.csv")
```


```{r}
# Landscape covariates
covariates <- read_csv("./data/processed/OSM_site_covariates_2021-2024.csv") %>%
  # Add a column from the spreadsheet of OSM land-use classifications)

  left_join(read_csv("./data/raw/landscape_unit_land_use_classifications.csv", 
                     by = "array"))
```

Quick and dirty fix to ensure we only plot sites for which we have detection data. Handles missing data from image tagging that we should track down later but maybe after FYE. If we constrain this to sites contained within "independent detections" file we'll be able to filter everything to only the sites for which we actually have data. 

```{r}
sites <- (unique(indet$site))
```

## 1.2. Display camera operability 

Let's read in the camera operability dates *for cameras with data* and plot it out. For now, we'll focus this entire script on just the 2024-2025 data since that's what we want for the Fiscal report. Importing the *entire* dataset means we can reuse this code for the data paper, or project-wide summaries as needed without major revision. 


Now we can plot it:

```{r}
operating %>%
  
    # This query should pull out only the 2024-2025 data from these arrays :)
  filter(year >= 2024,
         array %in% c("LU4", "LU2", "LU3", "LU8")) %>%

  filter(site %in% sites) %>%
  
  ggplot(., aes(x = date, y = site, fill = array)) +
    geom_tile() +
    scale_x_date(expand = c(0, 0)) +
    theme_classic()
```


## 1.3. Indpendent detections for mammals

Now lets create a few quick figures to look at the detection data.

First create a vector of the list of mammals to use for quick data visualization/exploration.

```{r graph mammal detections}
mammals <- c('White-tailed deer',
             'Black bear',
             'Snowshoe hare',
             'Moose',
             'Coyote',
             'Fisher',
             'Red squirrel',
             'Striped skunk',
             'Grey wolf',
             'Red fox',
             'Cougar',
             'Lynx',
             'Short-tailed weasel',
             'Porcupine',
             'Beaver',
             'Martin',
             'Wolverine',
             'Caribou',
             'Long-tailed weasel')
```

Now plot it: 

```{r graph mammal detections}
# remove NAs and select just images with mammals first then pipe new data into ggplot
det_graph <- indet %>% 
  
    # This query should pull out only the 2024-2025 data from these arrays :)
  filter(year >= 2024,
         array %in% c("LU4", "LU2", "LU3", "LU8")) %>%

  # remove less useful species
  filter(species %in% tolower(mammals)) %>% 
  
  # get the number of individual detections per species to add to graph
  group_by(species) %>% 
  
  mutate(n = n()) %>% 
  
  ungroup() %>% 
  
  ggplot(.,
         aes(x = species)) +
  
  # create bar graph of the counts of each spp in the data
  geom_bar(aes(fill = species)) +
  
  # add the number of detections above each bar using the variable n we calculated earlier
  geom_text(aes(label = n,
                y = n + 50),
            size = 4) +
  
  # change y axis label
  labs(y = 'Number of independent detections') +
  
  # change breaks for y axis
  scale_y_continuous(breaks = seq(0,3500, by = 250)) +
  
  # change theme elements
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1,
                                   size = 14),
        axis.title = element_text(size = 16),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank()) 

# print graph
det_graph

```

If we want to save the plot for easier viewing later we can use the code below

```{r save det_graph, eval=FALSE}

# save graph as jpeg (can also save as tiff, png, pdf by changing the file extension) but don't use .tiff in the github repo it takes up too much space and causes issues
ggsave('2024_indv_det_graph.jpeg',
       det_graph,
       path = 'figures',
       width = 12,
       height = 10,
       units = 'in',
       dpi = 600)

```

## 1.4. Independent detections per LU

let's also create one that graphs each LU in it's own panel using facet_wrap

```{r det graph LUs}
# let's also create one that graphs each LU in it's own panel using facet_wrap
det_plot_LUs <- indet %>% 
  
  # This query should pull out only the 2024-2025 data from these arrays :)
  filter(year >= 2024,
         array %in% c("LU4", "LU2", "LU3", "LU8")) %>%
  
  # remove less useful species
  filter(species %in% tolower(mammals)) %>% 
  
 # group by array and species to calculate dets per spp per LU
 group_by(array, species) %>% 
  
  # calculate a column with unique accounts of each species
  reframe(count = n_distinct(event_id)) %>% 
  
  # pipe to ggplot and set aesthetics mapping
  ggplot(aes(x = reorder(species, count), y = count)) +
  
  # plot as bar graph
  geom_col() +
  
  # plot each LU in own panel
  facet_wrap(vars(array)) +
  
  # add the number of detections at the end of each bar
  geom_text(aes(label = count),
            color = "black",
            size = 3,
            hjust = 0.2,
            vjust = -0.3) +
  
  # label x and y axis with informative titles
  labs(x = 'Species',
       y = 'Number of Independent (30 min) Detections') +
  
  # add title to plot with LU name
  
  #ggtitle("LU21 Detections")+
  
  # set the theme
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1,
                                   size = 12))

# view plot
det_plot_LUs
```

We can also save this plot if we want it for reports etc. 

```{r, eval = FALSE}
# save this plot

ggsave('figures/OSM_ind_det_per_LU_2023.jpg',
       det_plot_LUs,
       dpi = 600,
       width = 10,
       height = 12,
       units = 'in')
```



## 1.5. Naive occupancy by species/array

Based on total detections what is the naive occupancy?

```{r fig.height=8, fig.width=10}
# Naive occupancy
fig_naiveocc <- totaldet %>%
  
  group_by(species, array) %>%
  
  summarize(presence = sum(presence), 
            total = n(),
            naive_occ = presence/total) %>%
  
  ggplot(., aes(x = array)) + 
  
    geom_bar(aes(y = naive_occ), stat = "identity", fill = "grey50") +  # Total sites in grey
  
    #geom_bar(aes(y = n_present), stat = "identity", fill = "firebrick") +  # Present sites in red
  
    geom_text(aes(y = naive_occ + 0.1, label = round(naive_occ, 2)), size = 2.7, fontface = "italic") +  # Label above bars
  
    theme_bw() + 
  
    facet_wrap(~species, scales = 'fixed') +
  
    labs(x = "array",
         y = "naive occupancy") + 
    

    theme(
         axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
         strip.background = element_blank(),  # Remove facet background
         strip.text = element_text(face = "bold", size = 10)
         ) + 
    
    scale_y_continuous(expand = expansion(mult = c(0, 0.05)))  # Remove bottom space, keep small top space

fig_naiveocc
```

## 1.6. Monthly presence:

Let's also try to plot the presence data at least for each species so we can see which species we likely have enough data for to model

```{r plot prop month dets}

# first reformat the data so species presence is a column
proportional_detection_plot <- monthlyprop %>%
  
  pivot_longer(cols = c('black_bear':'wolverine'),
               names_to = 'species_presence',
               values_to = 'months_present') %>% 
  
  select(site, array, species_presence, months_present) %>% 
  
  ggplot(aes(x = site, y = months_present, fill = array)) +
  
  # plot as bars
  geom_col() +
  
  # viridis for colorblind :)
  scale_fill_viridis_d() +
  
  labs(y = "Months present", 
       x = "Site") + 
  
  scale_y_continuous(breaks = seq(0, 14, 2),
                     expand = c(0, 0)) +
  
  # use facet wrap to generate separate plots for each species
  facet_wrap(vars(species_presence)) + 
  
  theme_bw() + 
  
  theme(axis.text.x = element_blank(),
        panel.grid = element_blank())  # remove x-axis labels

# view plot
proportional_detection_plot
```

If we want to save the plot for easier viewing later we can use the code below

```{r save plot prop month dets, eval = FALSE}

# save graph as jpeg (can also save as tiff, png, pdf by changing the file extension)
ggsave('OSM_proportional_detections_2024.jpeg',
       proportional_detection_plot,
       path = 'figures',
       width = 12,
       height = 8,
       units = 'in',
       dpi = 600)
```



********* NEEDS TO BE ADDED FOR FYE DEADLINE

## 1.7. Change in naive occupancy for LU2 and LU3 (MARISSA)

Will need to re-import data, or just move the `filter(year>=2024)` etc. queries to the individual figures above so the full data is present. 

```{r}

```



# 2. Covariate figures

## 2.1. Covariate histogram by LU

First, well pads: 

```{r}
covariates %>%
  
  # Remove the early, partial sample of LU2 and LU3 (the only arrays from 2022)
  filter(array_year > 2022,
         # Buffer 500m is the geometric mean of our sampled scales, let's use it. 
         buffer_dist == 500) %>%
  
  ggplot., 
```

