---
title: "Untitled"
author: "Aidan Brushett"
date: "2025-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

OLD CODE


## Graphs

### Indpendent detections for mammals

Now lets create a few quick figures to look at the detection data

```{r graph mammal detections}

# Data visualization independent detections---------------------------------------------

# read in saved detection data if starting here
detections <- read_csv('data/processed/OSM_ind_det_2024.csv') %>% 
  
  # change site, species and event_id to factor
  mutate_if(is.character,
            as.factor)

# check number of different species
levels(OSM_2024_det_ind$species)

# create a vector of the list of mammals to use for quick data visualization/exploration. Could also create a list of species that aren't useful if that is shorter and use filter(!species %in% OBJECTNAME) but for this example the vectors were about the same length
mammals <- c('White-tailed deer',
             'Black bear',
             'Snowshoe hare',
             'Moose',
             'Coyote',
             'Fisher',
             'Red squirrel',
             'Striped skunk',
             'Grey wolf',
             'Red fox',
             'Cougar',
             'Lynx',
             'Short-tailed weasel',
             'Porcupine',
             'Beaver',
             'Martin',
             'Wolverine',
             'Caribou',
             'Long-tailed weasel')

# remove NAs and select just images with mammals first then pipe new data into ggplot
det_graph <- detections %>% 
  
  # remove less useful species
  filter(species %in% mammals) %>% 
  
  # get the number of individual detections per species to add to graph
  group_by(species) %>% 
  
  mutate(n = n()) %>% 
  
  ungroup() %>% 
  
  ggplot(.,
         aes(x = species)) +
  
  # create bar graph of the counts of each spp in the data
  geom_bar(aes(fill = species)) +
  
  # add the number of detections above each bar using the variable n we calculated earlier
  geom_text(aes(label = n,
                y = n + 50),
            size = 4) +
  
  # change y axis label
  labs(y = 'Number of independent detections') +
  
  # change breaks for y axis
  scale_y_continuous(breaks = seq(0,3500, by = 250)) +
  
  # change theme elements
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1,
                                   size = 14),
        axis.title = element_text(size = 16),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank()) 

# print graph
det_graph

```

If we want to save the plot for easier viewing later we can use the code below

```{r save det_graph, eval=FALSE}

# save graph as jpeg (can also save as tiff, png, pdf by changing the file extension) but don't use .tiff in the github repo it takes up too much space and causes issues
ggsave('2024_indv_det_graph.jpeg',
       det_graph,
       path = 'figures',
       width = 12,
       height = 10,
       units = 'in',
       dpi = 600)

```

### Independent detections per LU

let's also create one that graphs each LU in it's own panel using facet_wrap

```{r det graph LUs}
# let's also create one that graphs each LU in it's own panel using facet_wrap
det_plot_LUs <- detections %>% 
  
  # remove less useful species
  filter(species %in% mammals) %>% 
  
 # group by array and species to calculate dets per spp per LU
 group_by(array, species) %>% 
  
  # calculate a column with unique accounts of each species
  reframe(count = n_distinct(event_id)) %>% 
  
  # pipe to ggplot and set aesthetics mapping
  ggplot(aes(x = reorder(species, count), y = count)) +
  
  # plot as bar graph
  geom_col() +
  
  # plot each LU in own panel
  facet_wrap(vars(array)) +
  
  # add the number of detections at the end of each bar
  geom_text(aes(label = count),
            color = "black",
            size = 3,
            hjust = 0.2,
            vjust = -0.3) +
  
  # label x and y axis with informative titles
  labs(x = 'Species',
       y = 'Number of Independent (30 min) Detections') +
  
  # add title to plot with LU name
  
  #ggtitle("LU21 Detections")+
  
  # set the theme
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1,
                                   size = 12))

# view plot
det_plot_LUs
```

We can also save this plot if we want it for reports etc. 

```{r, eval = FALSE}
# save this plot

ggsave('figures/OSM_ind_det_per_LU_2023.jpg',
       det_plot_LUs,
       dpi = 600,
       width = 10,
       height = 12,
       units = 'in')
```

# Response metrics

there are several response metrics we can calculate, the ones we will cover here are. 

 1. Total independent detections per species/site
 2. Presence/absence per species/site
 3. Proportion of monthly detections

> Generally we only need #3 (proportional monthly detections) but we provided data for ES 482/582 class and wanted them to have multiple responses metrics to choose from for modeling purposes. 

## Data

For this we need the deployment and independent detection data we created earlier, if you are still working through this script its the 'deploy_fixed' & 'detections' objects

```{r response metric data}
# deploy
deploy <- read_csv('data/raw/OSM_deployment_2024.csv')


# detections
detections <- read_csv('data/processed/OSM_ind_det_2024.csv') %>% 
  
  # change site, species and event_id to factor
  mutate_if(is.character,
            as.factor)
```

For plotting and formatting proportional monthly detections we need to create a subset of the species in the detections data to just include several focal species we are interested in
```{r focal species}

# create a list of focal species for filtering the data/plots
 focal_species <- c('Black bear',
                    'Caribou',
                    'Cougar',
                    'Coyote',
                    'Fisher',
                    'Grey wolf',
                    'Lynx',
                    'Moose',
                    'Red fox',
                    'White-tailed deer',
                    'Wolverine',
                    'Snowshoe hare', # this one was added for ES 482/582 class data, we don't normally use for OSM reports
                    'Red squirrel' # this one was added for ES 482/582 class data, we don't normally use for OSM reports
                    )
```


## 1. Total independent detections 

The first response metric we will calculate is the total number of independent detections per species per site. 

To do this we use the detections data we created earlier from the raw Timelapse data. We need to group by site and species and then we can use the `summarise()` function with the `n()` function to count the total detections.

After that we ungroup the data so if we run an analysis or make a plot it doesn't stay grouped and then we use the `pivot_wider()` function to make a column for each species and a row for each site.

Finally we need to replace any NAs with zeros (the `n()` function function won't insert a zero if there aren't any observations to count so these NAs are indeed zeros)
```{r total ind det}

total_detections <- detections %>% 
  
  # group by site and species to count detections
  group_by(site, species) %>% 
  
  # use summarise to count detections per species per site
  summarise(detections = n()) %>% 
  
  ungroup() %>% 
  
  pivot_wider(names_from = species,
              values_from = detections) %>% 
  
  # replace NAs with 0 in all species columns
  mutate(across(
    where(is.numeric),
    ~ replace_na(., 0)))
  
```

Now that this data is formatted we should save it to the data/processed folder for use later

```{r save total ind det}

# save csv file to processed data folder 
write_csv(total_detections,
          'data/processed/OSM_total_detections_2024.csv')
```

We can also plot this data to see what it looks like

In the code below we create an object called site_detections_plot where we pipe the total detection data into the `ggplot()` function after doing some formatting to make it plot

```{r plot total ind det}

site_detections_plot <-  total_detections %>% 
  
  # we need to pivot longer to create species column again for plotting
  pivot_longer(cols = 2:37,
               names_to = 'species',
               values_to = 'detections') %>% 
  
  # remove less useful species using a list created in the 'Graph independent detections' section
  filter(species %in% mammals) %>% 
  
  # pipe into ggplot function
ggplot(.,
       aes(x = site,
           y = detections)) +
  
  geom_col() +
  
  # use facet wrap to make separate plots for each species 
  facet_wrap(vars(species)) +
  
  # shift axis text to 90 degrees so site name are readable
  theme(axis.text.x = element_text(angle = 90,
                                   size = 3),
        axis.ticks.x = element_blank())


# view plot
site_detections_plot

```

This is not the most readable plot because some species are skewing the x axis really high but it works for exploratory purposes.

If we want to save the plot for easier viewing later we can use the code below

```{r save plot total ind det, eval=FALSE}

# save graph as jpeg (can also save as tiff, png, pdf by changing the file extension)
ggsave('OSM_total_detections_site_2024.jpeg',
       site_detections_plot,
       path = 'figures',
       width = 12,
       height = 10,
       units = 'in',
       dpi = 600)

```


## 2. Presence/absences

A second response metric we may want to use is simply presence/absence data. Here we can use the total_detections data and replace any values greater than 0 with 1s to create a binary variable. 

```{r presence/absence}
# we can use the data from above to create a similar response metric of simply presences and absences per species per site

species_presence <- total_detections %>% 
  
  # replace all values above 0 with 1s
  mutate_if(is.numeric, 
            ~1 * (. > 0))

# now we have presence absence data for all species
```

We can save this to data/processed

```{r save presence/absences}
# save csv file to processed data folder 
write_csv(species_presence,
          'data/processed/OSM_total_presence_absence_2024.csv')
```

We may also want to plot this similarly to the total detections per site to look at the data easier

As before we create a new object with our plot title name and then pipe the data after some formatting into the `ggplot()` function.
```{r plot presence/absence}
# plot this data in ggplot
species_presence_plot <- species_presence %>% 
  
  # first we need to pivot the data longer again so we have a species column for plotting
  pivot_longer(cols = 2:37,
               names_to = 'species',
               values_to = 'presence') %>% 
  
  # remove less useful species using a list created in the 'Graph independent detections' section
  filter(species %in% mammals) %>% 
  
  # pipe into ggplot function
ggplot(., 
       aes(x = site, y = presence)) +
  
  # use geom_jitter instead of geom_point so we can shift points on y-axis to make them easier to view
  geom_jitter(shape = 16,
              size = 1.5,
              width = 0,
              height = 0.05,
              alpha = 0.5) +
  
  # use facet_wrap to make separate plots for each species so data is easier to look at
  facet_wrap(vars(species)) +
  
  # shift axis text to 90 degrees so site name are readable
  theme(axis.text.x = element_text(angle = 90,
                                   size = 3))

# view plot
species_presence_plot
```

If we want to save the plot for easier viewing later we can use the code below

```{r save plot presence/absence, eval=FALSE}

# save graph as jpeg (can also save as tiff, png, pdf by changing the file extension)
ggsave('OSM_total_presence_absence_2024.jpeg',
       species_presence_plot,
       path = 'figures',
       width = 12,
       height = 10,
       units = 'in',
       dpi = 600)
```

## 3. Proportion monthly detections

We need to use the deployment data to determine how many days each camera was active for

First we create a new data frame from the deploy_fixed (deployment data frame that's been cleaned) with some of the same columns and a new column 'day' that goes from the start date to the end date of each camera deployment and increases by intervals of 1 (for each day the camera was active). We use this to calculate the number of days/month each camera was active for and create a new variable for this called days_month. and then we create another column based on the days/month the camera was active that classifies each month of data for a camera as keep or remove based on whether there were at least 15 active camera days in that month. This is because we don't want to estimate monthly data for a camera that was working less than half the time. 

```{r prop month detections step 1: days active}
deploy_active <- deploy %>%
  
  drop_na(start_date, end_date) %>% # remove NA rows. There is one, from a camera that was not retrieved :(
  
  rowwise() %>%
  
  # create a list sequence of dates for each site from the first to the last day of operability
  mutate(date = list(seq(from = date(start_date), to = date(end_date), by="day"))) %>% 
    
  unnest(date) %>%
  
  mutate(year = year(date),
         month = month(date))
```

We also need to modify the OSM operating dataframe to account for days where the camera was fully covered by snow or otherwise obstructed (cases where `cameramalfunction == *'Fully obscured'*`). Let's identify those from the raw timelapse data and remove them from `osm_operating`.

```{r}
deploy_obscured <- OSM_2023_data %>%
  
  filter(cameramalfunction == 'Fully obscured') %>%
  
  mutate(date = as.Date(datetime)) %>%
  
  select(array, site, date) %>% 
  
  distinct()
```

Then, remove those dates from `deploy_active`. 

```{r}
deploy_active <- deploy_active %>%
  
  anti_join(deploy_obscured)
```

Now we calculate the total number of months each camera was active for. We will impose an additional restraint that 'active' months must have been operating for at least 15 days/month. We will use this data again later.

```{r prop month detections step 2: months active}
# calculate the total number of months each camera was active for including only those active for >15 days/month or the 'keep' values
deploy_active_months <- deploy_active %>%

  group_by(array, site, year, month) %>%
  
  summarize(days_active = n()) %>%
  
  filter(days_active >= 15) %>% # retain months where there are >15 days of data only.

  select(-days_active)
# we will use this data later

```

Now that we have identified cameras that were not active long enough each month to reliably extract data from. Let's merge our detection data into this dataframe of camera operability.

First, organize our independent detections into monthly bins. These will be the months of true presence (presence==1) in our final dataset. 

```{r}
monthly_presence <- detections %>%
  
  distinct(array, site, species, month, year) %>% # keep one copy of detections for each month/year
  
#  filter(month %in% all_operating_months$month) %>% # retain only the data for weeks with sufficient number of days >4. The rest we don't want. 
  mutate(presence = 1)
```

Second, let's merge our monthly presence into the list of *all* operating months with >15 days. Where NAs are introduced into the *presence* column, there was no detection of the given species in that month. We will replace these NAs with "0" to denote 'true' absence. 

```{r}
monthly_presenceabsence <- deploy_active_months %>%
  
  crossing(species = unique(as.character(monthly_presence$species))) %>%
  
  left_join(monthly_presence, by = c("array", "site", "month", "species", "year")) %>%
  
  mutate(presence = replace_na(presence, 0), # Fill missing detections with 0
         species = as.factor(species))
```

Save the presence-absence dataset:

```{r, eval = FALSE}
write_csv(monthly_presenceabsence,
          './data/processed/OSM_monthly_presence_absence_2024.csv')
```

Then from the data we keep we can create a new data frame that has 1 row per camera and a column for each species indicating how many of the active months each species was detected at that camera.

```{r prop month detections step 3: presences}

# now that we have  identified cameras that were not active long enough each month to reliably extract data from we can use that column to remove this data from the detections data frame

proportional_detections <- monthly_presenceabsence %>% 
  
  # group by site and species to create data frame with one row per site x species combo
  group_by(site, array, species) %>% 
  
  summarise(
    months_present = sum(presence > 0),
    months_active = n(),
    .groups = "drop"
  ) %>%

  # filter to only species in the list of focal species we created earlier
  # NOTE when we do this we lose 3 sites because there weren't any of these species detected at those sites during months where the camera was active >= 15 days/month
  filter(species %in% focal_species) %>% 
  
  # pivot the data wider so there is a column for each species and 1 row per site
  pivot_wider(names_from = species,
              values_from = months_present) %>%

  # replace NAs with zeros in all species columns
  mutate(across(
    where(is.numeric),
    ~ replace_na(., 0))) %>% 

  # ensure all species columns are numeric not integer
  mutate_if(is.integer,
            as.numeric)
```

Now we can run the function below to create a second column for each species that will represent the number of absences (months the species was not detected) at each camera from the active months.

```{r prop month detections step 4: absences}

# run a function to create columns for absences based on presence data and how many months the camera was functioning

# first convert data to data frame not a tibble for function to work
proportional_detections <- as.data.frame(proportional_detections)

# create a vector of the species columns for the loop
# use all species columns (this value may change year to year)
cols <- 4:16

for (col in cols) {
  if (is.numeric(proportional_detections[,col]) & is.numeric(proportional_detections$months_active)) 
    {new_col_name <- paste0("absent_", colnames(proportional_detections)[col])
    proportional_detections[new_col_name] <- proportional_detections$months_active - proportional_detections[,col]
  }
}

```

We want to rename the species columns so they don't have spaces (r doesn't like spaces for column names but it was fine when they were entries in the data)

```{r prop month detections step 4: rename species columns}

# rename columns for species because R doesn't like spaces
proportional_detections <- proportional_detections %>% 
  
  # set the column names to lower case and replace the spaces with '_' (these are both personal preferences of mine)
  set_names(
    names(.) %>% 
      tolower()%>% 
      str_replace_all(pattern = ' ',
                      replacement = '_'))

```

Now we have to do a bit of final wrangling of the data to fix the bear columns because we don't want to consider the months that bears are not active in the months active columns. 

```{r prop month detections step 5: bear fix}

# fix bear data

# before we can use this data we need to adjust the columns for bears since they are hibernating we don't want to calculate their presence/absence for those inactive months

# now let's recalculate the number of active months
months_active_bears <- deploy_active_months %>% 
  
  # filter to months bears are active (April - November)
  dplyr::filter(month %in% c("4", "5", "6", "7", "8", "9", "10", "11")) %>% 
  
  # get distinct rows for each 
  distinct(site, month, year, 
           .keep_all = TRUE) %>% 
  
  # group by site
  group_by(site) %>% 
  
  # count the number of months active during bear active season and save as new column
  summarise(months_active_bears = n())
 

# now we overwrite the absent column for black bears using new info

proportional_detections_bears <- proportional_detections %>% 
  
  # join the bear active data
  left_join(months_active_bears, 
            by = 'site') %>% 
  
  # overwrite absent black bear column
  mutate(absent_black_bear = months_active_bears - black_bear) %>%
  
  # get rid of unnecessary columns for active months
  select(-c(months_active, 
            months_active_bears))

```

Finally we can save this data
```{r save prop month detections}
# save data
write_csv(proportional_detections_bears,
          'data/processed/OSM_proportional_detections_2024.csv')
```

Let's also try to plot the presence data at least for each species so we can see which species we likely have enough data for to model

```{r plot prop month dets}

# first reformat the data so species presence is a column
proportional_detection_plot <- proportional_detections_bears %>% 
  
  pivot_longer(cols = c('black_bear':'wolverine'),
               names_to = 'species_presence',
               values_to = 'months_present') %>% 
  
  select(site, array, species_presence, months_present) %>% 
  
  ggplot(aes(x = site, y = months_present, fill = array)) +
  
  # plot as bars
  geom_col() +
  
  # viridis for colorblind :)
  scale_fill_viridis_d() +
  
  # use facet wrap to generate separate plots for each species
  facet_wrap(vars(species_presence))

# view plot
proportional_detection_plot
```

If we want to save the plot for easier viewing later we can use the code below

```{r save plot prop month dets, eval = FALSE}

# save graph as jpeg (can also save as tiff, png, pdf by changing the file extension)
ggsave('OSM_proportional_detections_2024.jpeg',
       proportional_detection_plot,
       path = 'figures',
       width = 12,
       height = 10,
       units = 'in',
       dpi = 600)
```
